### 2018-02-04 paper/proposal.tex
first

\documentclass{article}

\usepackage{geometry}
\geometry{margin=1in}

\usepackage{graphicx}
\graphicspath{ {images/}}

\usepackage[backend=biber,style=authoryear]{biblatex}
\addbibresource{proposal.bib}

\author{Tom Wallace}

\title{Paper Proposal 
\\ \large The Rise of Slavery in America: an Agent-Based Model}

\begin{document}

\maketitle

What caused chattel slavery to emerge in the American South? This question has generated an enormous amount of
academic inquiry. Most agree that American slavery had both ideological and
economic underpinnings, but the relative importance of these two factors is
contested. The ideology camp sees slavery as driven by white supremacy: a 
``peculiar institution'' (to use John Calhoun's phrase) 
that was carried out for mostly non-economic reasons. In this view,
white Southerners believed in the inferiority of blacks and so erected an economic
system actualizing those beliefs; whether it was profitable or not was a
secondary concern. The ``new history of capitalism'' (NHC) 
emphasizes the profit-motivated nature of slavery and 
in particular its connection to the emergence of industrial capitalism
and global trade in the 19th century. In this view, white supremacist ideology was constructed as 
\textit{post hoc} intellectual justification for a money-making scheme predicated
on cheap black labor and high cotton prices. 
To simplify the debate, did racist ideology
dictate the economic system, or did the economic system dictate racist 
ideology?

Sven Beckert's \textit{Empire of Cotton}, an influential entry in the
NHC literature, offers a compelling hypothesis centered on market information,
labor mobility, and land use. The Industrial Revolution caused great demand
for foreign cotton to supply British textile mills. However, extant cotton production
in the Global South was governed more by local culture and traditions than
capitalist logic. To paraphrase, peasants in then-British India 
grew some cotton on the side because that was what they had always done, 
but they were unwilling to radically ramp up production (which would have
required abandonment of food crops and organization into large plantations rather 
than family plots) to meet the demands of Lancashire and Birmingham factories. 
There was a need for a industrial-scale cotton production system more attuned to 
global price signals.

Beckert's hypothesis is that American slavery resulted from 
land-use and labor mobility issues associated with the creation
of such a system. Cotton is nutrient-hungry (it rapidly exhausts the soil) and labor-intensive
(land-clearing and cultivation is hard work). From the perspective of British ``cotton
capitalists'' who would eventually finance virtually all U.S. plantations, 
an ideal cotton production set-up would be highly sensitive to price signals; 
surplus land to which production could move after local soil exhaustion; 
no competition with food crops or land rights; and a highly mobile workforce 
that would follow the rapid movement of cotton farms from exhausted to
fresh soil. The early 19th century American South had almost all these
characteristics. The climate was favorable; the culture was mercantile; 
advances in communications technology (e.g. the telegraph) allowed the ever-more-rapid transmission of 
market information; the Louisiana Purchase and depopulation of Native Americans 
left a vast frontier of virgin soil.

The only barriers were land use and labor. Free white
men had property rights and refused to totally structure their land use 
around cotton (whose price was volatile, making agricultural 
diversification a better strategy for small farmers without a large financial cushion).
They also were unwilling to pack up and move every few years to follow the path
of cotton cultivation. Slavery is argued to
have evolved as a response to these land use and labor mobility issues.
By importing Africans and stripping them of all rights, 
one could create a labor force that did not have a choice 
about how to use the land or where to move, and so could be coerced into
profit-maximizing patterns (for example, the modern geographic distribution of 
African-Americans still matches the path of cotton expansion in the
1800s, as seen in Figures 1 and 2). In essence, the NHC hypothesis is that when 
market information was local, there was little incentive
for slavery (and indeed, slavery had a surprisingly small footprint in America
in the 16-18th centuries, at least compared to its later heights); 
when markets became global, white agents independently converged on
slavery as a mechanism for  matching cotton supply in Mississippi to cotton demand in Lancashire.

\begin{figure}[h!]
\caption{Historical distribution of cotton production}
\includegraphics[scale=0.75]{historical}
\centering
\end{figure}

\begin{figure}[h!]
\caption{Current distribution of African-American population}
\includegraphics[scale=0.4]{current}
\centering
\end{figure}

Many aspects of the NHC hypothesis are amenable to spatial 
agent-based modeling. The NHC hypothesis relies on autonomous agents reacting to
information with boundedly rational behavior. It emphasizes the
interaction of agents with the spatial environment in terms of labor mobility
(where in the environment agents chose to go) and land-use (how they shape the
environment once there). It consists of different levels of analysis, including
both the individual and system. It emphasizes the effect of information
availability (local vs. global). These features are well-suited for agent-based
modeling.

I propose a spatial agent-based model. The environment consists of grid-square
cells. Human agents occupy these cells. Agents must decide between two land use
strategies: cotton-only or mixed cultivation. The former depletes the soil of
the grid-square while the latter can continue indefinitely. Once the soil is
depleted, the agent must move to a new grid-square. Agent cognition is governed
by preferences regarding land-use and labor mobility, and by profit motivation
based on the price of cotton and of mixed agriculture. Information regarding
prices is geographically bounded (I have not yet determined an exact mechanism
for this). The hypothesis is that increased information regarding prices causes
greater marginal value to labor mobility and land use. Clearly I have not
thought through all aspects of the model, but this is the general vision. I
likely would code this model in Python. The model would proceed in discrete
time-steps. In terms of data, I am considering obtaining GIS files of the
American South (though I do not think they are strictly necessary to explore the
hypothesis). The key dynamics that I want to model are:
\begin{itemize}
	\item Land use
	\item Labor mobility
	\item Information availability
	\item Global prices
\end{itemize}

\nocite{*}

\medskip
\printbibliography[title={}]

\end{document}

### 2018-04-24 prez/prez.tex
initial

\documentclass{beamer}

\usetheme{Madrid}
\usecolortheme{spruce}

\usepackage{graphicx}
\graphicspath{{./img/}}

\title{Management Application of an Agent-Based Model: Control of Cowbirds at
the Landscape Level}
\subtitle{Harper, Westervelt, and Shapiro (2002)}
\author{Tom Wallace}
\institute{George Mason University}
\date{Spring 2018}

\begin{document}

\frame{\titlepage}

\begin{frame}
	\frametitle{The greatest threats to the United States?}
	\includegraphics[height=1.5in, keepaspectratio]{fire.jpg}
	\includegraphics[height=1.5in, keepaspectratio]{terrorists.jpg}
	\includegraphics[height=1.5in, keepaspectratio]{collapse.jpeg}
\end{frame}

\begin{frame}
	\frametitle{The hidden menace}
	\begin{center}
		\includegraphics[height=2.5in, keepaspectratio]{bird.jpg}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{A bloody campaign of deception and sabotage}
	\begin{center}
		\includegraphics[height=1.5in, keepaspectratio]{nest.jpg}
		\includegraphics[height=1.5in, keepaspectratio]{trend.png}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Taking aim at the heart of American military power}
	\begin{center}
		\includegraphics[height=1.5in, keepaspectratio]{fort.jpg}
		\includegraphics[height=1.5in, keepaspectratio]{map.png}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Only the A(BM)-Team can stop this reign of terror}
		Authors
			\begin{itemize}
				\item \small Steven Harper, University of Georgia
				\item \small James Westervelt, U.S. Army Corps of Engineers
				\item \small Ann-Marie Shapiro, University of Illinois
					\\~\\
			\end{itemize}
		Other research
			\begin{itemize}
				\item \small Desert tortoises at Fort Irwin
				\item \small Endangered birds at Fort Hood \\~\\
			\end{itemize}

		Goals
			\begin{itemize}
				\item \small Build model to predict feeding locations of cowbirds
				\item \small Extend model to simulate trapping of cowbirds
				\item \small Use this capability to develop optimal trapping strategy
				\item \small Terminate with extreme prejudice
			\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Overview}
	GIS (GRASS) and ABM (SWARM) \\~\\

	58 km x 58 km (335,000 ha) area \\~\\

	1 day time step, 100 steps \\~\\

	Three agent types
	\begin{itemize}
		\item \small Feeding area (fixed)
		\item \small Cattle herd (mobile)
		\item \small Cowbird (mobile) \\~\\
	\end{itemize}

	Feeding area $\to$ cattle movement $\to$ bird movement $\to$ trap placement

\end{frame}

\begin{frame}
	\frametitle{Feeding areas}
	56 ha square \\~\\

	Main attribute of interest: \textbf{grazing suitability} \\~\\

	(Partially) determines propensity of cow herd to move to that square \\~\\

	Determined by four sub-attributes
	\begin{itemize}
		\item \small Amount of grassland
		\item \small Patchiness of grassland
		\item \small Distance from corral
		\item \small Distance from permanent water \\~\\
	\end{itemize}
	\begin{center}
		\includegraphics[height=1.0in, width=4in]{grassland.jpg}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Cattle herds}

	30 head per herd, 464 herds, initialized at 116 corrals \\~\\

	Move around environment grazing according to ``win-switch'' strategy\\~\\

	Movement logic: pick cell with highest \textbf{grazing quality}
	\begin{itemize}
		\item \small Grazing suitability (per previous slide)
		\item \small Time since previous occupation
		\item \small Distance from current location \\~\\
	\end{itemize}

	Move after 8 days

	\begin{center}
		\includegraphics[height=1.0in, keepaspectratio]{herd.jpg}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Cowbirds}

	Randomly initialized in breeding-area squares \\~\\

	Cognition for moving from breeding area to feeding area\\~\\

	Four possible movement rules 
	\begin{itemize}
		\item \small Next-nearest: move randomly to nearest feeding area until cows found
		\item \small Memory: try places have been before
		\item \small Memory-with-perception: add ability to assess feeding areas en-route
		\item \small Omniscient: benchmark\\~\\
	\end{itemize}
	\begin{center}
		\includegraphics[height=1.0in, keepaspectratio]{bird.jpg}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Trapping}
	Two new agent types \\~\\

	\textbf{Traps} 
	\begin{itemize}
		\item \small Can be mobile or fixed
		\item \small Occupy cell
		\item \small Have some probability of capturing a bird that visits cell \\~\\
	\end{itemize}

	\textbf{Trap managers}
	\begin{itemize}
		\item \small Goal: place traps where most cowbirds are
		\item \small Places traps and moves mobile ones
		\item \small Cognition for doing so is under-described
	\end{itemize}

	\begin{center}
		\includegraphics[height=1.0in, keepaspectratio]{trap.jpeg}
	\end{center}

\end{frame}

\begin{frame}
	\frametitle{V\&V and results}

	Parameter values well-justified and subjected to sensitivity analyses
	\begin{itemize}
		\item \small Field studies, literature review, etc.
		\item \small Many different runs with different settings (e.g. cowbird cognition) \\~\\
	\end{itemize}

	Simulation appears to qualitatively match real life \\~\\

	Fixed traps more effective (entirely based on parameter $p(\mathrm{catch)})$\\~\\

	Model-based trapping strategy caught 42\% more than real life
\end{frame}

\begin{frame}
	\frametitle{Model effectiveness and course themes}

	In our taxonomy: analyzed-analyzed \\~\\

	In my opinion, one of the best models this semester
	\begin{itemize}
		\item \small Clear, scoped purpose (no \textit{exploration of potential} here!)
		\item \small Logical and well-justified modeling choices
		\item \small Solves real-world problem \\~\\
	\end{itemize}

	Somewhat depressing that best models (e.g. this and Yellowstone elk) are of animals with much more limited
	behavior space than humans \\~\\

	Course themes:
	\begin{itemize}
		\item \small ABM-GIS integration: serves as a great example of a problem that is nearly-intractable
			unless you include both elements and integrate
	\end{itemize}
\end{frame}

\end{document}

### 2018-04-27 paper2/paper2.tex
werk

\documentclass{article}

\title{An agent-based model of military mechanization}

\usepackage[citestyle=authoryear, style=authoryear]{biblatex}
\addbibresource{paper2.bib}

\author{Tom Wallace}

\begin{document}

\maketitle

\section{Introduction}

This paper presents an agent-based model of military mechanization. First, it
provides a theoretical discussion of military mechanization, states the research
hypothesis, and explains why agent-based modeling is an appropriate research strategy.
Then, it details an original agent-based model; presents results; and discusses
their implications for the academic debate on military mechanization.

In summary, the orthodox

\section{Background}

States differ in the composition of their militaries, and in particular, of
their ground forces.\footnote{States may have multiple organizations that
conduct ground combat operations: e.g., the United States operates both an Army
and a Marine Corps as independent services. Henceforth, this paper uses the
terms ``army'' and ``ground forces'' interchangeably.} One notable axis of comparison 
is \textit{mechanization}: the degree to which the state relies on infantry 
vs. ground combat vehicles. Treating mechanization as a continuous variable, 
one extreme is an army consisting exclusively of dismounted, small 
arms-wielding infantry units; the other extreme is an army consisting 
solely of tank units, armored personnel carriers, artillery, and the like.
Clearly, no modern state embodies either extreme, but the 1970s-era Vietcong are
a good example of a low-mechanization army and the 1980s-era Israeli army of
high mechanization.

The consequences of choices regarding mechanization are extreme. To quote former
U.S. Secretary of Defense Donald Rumsfeld, ``you go to war with the army you
have.''

The causes of mechanization

\cite{sechser2010army}
\cite{biddle2004military}
\cite{gartzke2001democracy}
\cite{farrell2001transitional}

\section{Model}

\section{Results}

\section{Discussion}
\printbibliography[heading=bibnumbered]

\end{document}

### 2018-04-27 paper2/model.py

@@ -1,0 +1,40 @@
+def runSimulation():
+
+	years = [1979, 1981, 1983, 1985, 1987, 1989, 1991, 1993, 1995, 1997,
+			1999, 2001]
+
+	# get players for starting year (1979)
+	with open("players1979") as f:
+		agents1979 = f.readlines()
+	agents1979 = [x.strip() for x in agents1979]
+
+	# create agents
+	lst1979 = []
+	for name in agents1979:
+		agent = Agent(name)
+		lst1979.append(agent)
+
+	# read neighbors data
+	for agent in lst1979:
+		agent.addNeighbor("foo")
+		agent.addNeighbor("bar")
+		print(agent.neighbors)
+	
+class Agent:
+	def __init__(self, name):
+		self.name = name
+		self.log_mech = 0
+		self.neighbors = []
+		self.enemies = []
+
+	def addNeighbor(self, name):
+		self.neighbors.append(name)
+	
+
+'''
+lst = []
+for country in countries:
+	agent = State(country)
+	lst.append(agent)
+	'''
+runSimulation()

### 2018-04-28 paper2/model.py
foo

@@ -1,40 +1,119 @@
+'''
+Tom Wallace
+CSS 645
+Spring 2018
+Final project
+'''
+
+import numpy as np
+import pandas as pd
+import math
+
 def runSimulation():
 
-	years = [1979, 1981, 1983, 1985, 1987, 1989, 1991, 1993, 1995, 1997,
-			1999, 2001]
+	coeff_enemies = 1.2
+	coeff_neighbors = 1.15
+	coeff_aml = 0.9
+	coeff_em = 1.04 #differs from their actual coefficient
+	coeff_nm = 1.02 #differs from their actual coefficient
+	coeff_mountains = 1.1
+
+	# import mechanization data
+	sechser = pd.read_csv("sechser-saunders-ISQ-2010.tab", sep = '\t')
+
+	# import neighbors data
+	contdird = pd.read_csv("contdird.csv")
+
+	# import mid data
+	rivals = pd.read_csv("dyadmid20-1.csv")
+
+	# import cinc data
+	nmc = pd.read_csv("NMC_5_0.csv")
 
 	# get players for starting year (1979)
 	with open("players1979") as f:
 		agents1979 = f.readlines()
+	agents1979 = [x.replace('"', '') for x in agents1979]
 	agents1979 = [x.strip() for x in agents1979]
 
 	# create agents
-	lst1979 = []
+	lst = []
 	for name in agents1979:
 		agent = Agent(name)
-		lst1979.append(agent)
+		lst.append(agent)
 
-	# read neighbors data
-	for agent in lst1979:
-		agent.addNeighbor("foo")
-		agent.addNeighbor("bar")
-		print(agent.neighbors)
+	# get neighbors, enemies, aml, starting mech
+	for agent in lst:
+		agent.getNeighbors(1979, contdird)
+		agent.getEnemies(1979, rivals)
+		agent.getAML(1979, sechser)
+		agent.getStartingMech(1979, sechser)
+
+	# calculate enemy_mech and neighbor_mech
+	for agent in lst:
+		if(len(agent.enemies) > 0):
+			agent.calcEnemyMech(lst)
+			agent.calcNeighborMech(lst)
 	
+	# 1981
+	'''
+	for agent in lst:
+		print(agent.name, agent.current_mech, agent.mech, agent.neighbors, agent.neighbor_mech, agent.enemies,
+				agent.enemy_mech, agent.aml)
+	'''
+
 class Agent:
 	def __init__(self, name):
[…]

### 2018-05-06 paper2/paper2.tex
foo

@@
 \documentclass{article}
-
-\title{An agent-based model of military mechanization}
 
@@
 \addbibresource{paper2.bib}
+
+\title{An agent-based model of military mechanization}
 
@@
 
-In summary, the orthodox
+In summary, the orthodox view of military mechanization is that it is a
+reflexive process: states choose their mechanization posture largely based on
+that neighbors and enemies. This paper argues
+that this view---based on traditional longitudinal data analysis---is ripe for
+agentization:
 
@@
 
-States differ in the composition of their militaries, and in particular, of
-their ground forces.\footnote{States may have multiple organizations that
+\subsection{Military mechanization}
+
+States differ in the composition of their military ground forces.\footnote{States may have multiple organizations that
 conduct ground combat operations: e.g., the United States operates both an Army
@@
 terms ``army'' and ``ground forces'' interchangeably.} One notable axis of comparison 
-is \textit{mechanization}: the degree to which the state relies on infantry 
-vs. ground combat vehicles. Treating mechanization as a continuous variable, 
+is \textit{mechanization}: the degree to which an army consists of infantry 
+vs. combat vehicles. Treating mechanization as a continuous variable, 
 one extreme is an army consisting exclusively of dismounted, small 
@@
 
-The consequences of choices regarding mechanization are extreme. To quote former
-U.S. Secretary of Defense Donald Rumsfeld, ``you go to war with the army you
-have.''
+The consequences of choices regarding mechanization are large. There is
+academic consensus that defense policy---including but not limited to
+mechanization---conditions battlefield effectiveness. 
+\cite{lyall2009rage} finds that highly mechanized militaries are less effective
+at counterinsurgency, while \cite{biddle2004military} argues that
+mechanization-enabled mobility aids conventional warfighting. It thus
+pays to have an army that is designed to win the
+kinds of
+conflicts that a state is likely to face. Lacking a crystal ball, however,
+leaders cannot always forecast conflict, and armies are large, complex
+organizations that typically are held to change only infrequently and slowly at
+that, even if there is a pressing need (\cite{murray1998military},
+\cite{locher2004victory}, \cite{zegart2000flawed}). A state suddenly faced with
+a war for which its mechanization posture is ill-suited can either concede or
+pay a cost in blood and treasure; what it cannot do is
+quickly overcome the constraints imposed by past mechanization decisions. 
+A quote from then-Secretary of Defense Donald Rumsfeld---made in 2004, when the U.S. military
+was confronted in Iraq with the consequences of past choices regarding
+mechanization---emphasizes the point:
+``you go to war with the army you have'' (\cite{schmitt_2004}).
 
-The causes of mechanization
+The causes of military mechanization are disputed. \textit{Realists} believe 
+that mechanization choices are driven by security environment. States
+choose mechanization policies that they believe will allow them to prevail against potential adversaries, or to send a
+signal of deterrence, or to learn from the perceived mistakes of the last war
+(\cite{mearsheimer1983conventional};
+\cite{huth1988extended}; \cite{murray2011military}). The other school of thought is more
+diverse and so harder to name, but may be termed 
+\textit{institutionalists}. They believe that factors other than 
+strategic calculation determine military mechanization. Domestic political
+institutions---e.g., democracy vs. autocracy, politically stable vs.
+coup-prone, the state of civil-military relations, and so on---may influence 
+force structure (\cite{reiter2002democracies}; \cite{quinlivan1999coup};
+\cite{talmadge2015dictator}; \cite{brooks2008shaping}). Economic factor
+endowments may play a role: e.g., capital-rich states may mechanize more
+than labor-rich states (\cite{gartzke2001democracy}). Lastly,
+ideology (\cite{van1984cult}) and culture (e.g., \cite{pollack2004arabs}) are
+hard-to-measure but perhaps influential. Consider Ireland: between the 1920s and
+1940s, it expended a large portion of its defense budget on building
+a small number of tanks. They were wholly insufficient to repel a British 
+invasion (their ostensible purpose), and
+Ireland would have been better-served to pursue a low-mechanization guerrilla
+defense strategy, but tanks were seen as a prestigious signifier of a ``real''
+professional military, an important factor for the young Irish state
+(\cite{farrell1998professionalization};
+\cite{farrell2001transnational}).
 
-\cite{sechser2010army}
-\cite{biddle2004military}
-\cite{gartzke2001democracy}
-\cite{farrell2001transitional}
+\cite{sechser2010army} is the landmark study on military mechanization. They
+assemble a longitudinal dataset at the country-year level of the mechanization
+of all states between 1979 and 2001, and conduct regression analysis to identify
+the effect of covariates on mechanization. They conclude that ``choices about
+mechanization are strongly associated with a state's security environment.''
+The more mechanized a state's geographic neighbors and enemies, the higher the state's 
+own mechanization; also, states
+learn from defeat in insurgency by subsequently decreasing their mechanization.
+
+\subsection{The case for agent-based modeling}
+
+The findings of \cite{sechser2010army} are ripe for agentization. The
+conclusions drawn implicitly claim
+that states are autonomous agents; that they have
+cognition regarding how to perceive the outside world in terms of mechanization
+and how to adjust their own mechanization to it; that spatiality and
+neighborhood matter; that inter-agent relationships (e.g. whether you are my
+enemy or not) matter; and that
+agents learn from experience. Agent-based modeling is uniquely well-suited to
+represent such phenomena (\cite{gilbert2005simulation};
+\cite{miller2009complex}).
+
+Importantly, if the findings of \cite{sechser2010army} are correct and states
+largely determine their own mechanization levels on the basis of what other
+states are doing, a positive feedback loop exists. State A has high
+mechanization, causing enemy State B to raise their mechanization, causing State
+A to raise their mechanization... and so on. This feedback loop---often called
+the ``security dilemma''---has been extensively studied, including by Thomas Schelling, one of the pioneers
+of agent-based modeling (\cite{schelling1960strategy};
+\cite{schelling2006micromotives}).
+
+\subsection{Research question and strategy}
+
+I hypothesize that the arguments of \cite{sechser2010army} are incomplete and
+that some omitted variable or set of variables acts as a ``brake'' on military
+mechanization.\footnote{By ``omitted'', I mean that they conclude that the variable
+is not very important for mechanization, not that they do not address the
+variable at all.} An agent-based model is used to test this hypothesis. In broad
+terms, agents (states) are defined to have the cognitive and learning behavior
+implicitly claimed by \cite{sechser2010army}. The model is initialized with
+real-world data from 1979 and allowed to run until 2001. The micro- and
+macro-outcomes are then validated against real-world 2001 data. If
+\cite{sechser2010army} are correct, the model's 2001 outcomes should match
+real-world 2001 outcomes. If the model mechanization levels in 2001 are much
+higher, then their claimed behavioral rules are incomplete and fail to include
+some ``brake'' on mechanization, matching my hypothesis. Proving exactly what
+that missing variable or set of variables is is outside the scope of this model
+and paper, though it is discussed.
 
 \section{Model}
+
+This section of the paper describes the model. It does not use the ODD protocol
+of \cite{grimm2006standard} but covers many of the same points.
+
+\subsection{Agents}
+
+There is one agent type: states. Agents exist at a fixed location and do not
+move. The composition of agents changes throughout the
+
+\subsection{Environment}
+
+Agents exist in a spatial environment. Spatiality is represented by neighbor
+relationships rather than a raster or vector GIS map. Two states are defined as
+being neighbors if they share a contiguous land border or a stretch of water
+less than 400 miles. Every agent has an attribute \texttt{neighbor\_list}, a
+list in which it stores the names of neighboring agents for that time-step.
+
+\subsection{Initialization and Time}
+
+\subsection{Dynamics and Interactions}
 

### 2018-05-07 paper2/paper2.tex
foo

@@
 
+\begin{centering}
+	Word count: XXXX
+\end{centering}
 \section{Introduction}
@@
 
-Importantly, if the findings of \cite{sechser2010army} are correct and states
-largely determine their own mechanization levels on the basis of what other
-states are doing, a positive feedback loop exists. State A has high
+Importantly, the findings of \cite{sechser2010army} imply a positive feedback
+loop. State A has high
 mechanization, causing enemy State B to raise their mechanization, causing State
-A to raise their mechanization... and so on. This feedback loop---often called
-the ``security dilemma''---has been extensively studied, including by Thomas Schelling, one of the pioneers
+A to raise their mechanization... and so on. Such feedback loops are a hallmark
+of complex systems and again are well-suited for representation by agent-based
+modeling. This specific one---states changing their military in response to
+other states change in military in response to... and so on---feedback loop---is often called
+the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
 of agent-based modeling (\cite{schelling1960strategy};
@@
 There is one agent type: states. Agents exist at a fixed location and do not
-move. The composition of agents changes throughout the
+move. The set of agents encompasses all diplomatically-recognized 
+states in the world at a particular point in time, minus microstates (defined as
+those with a population less than 750,000).
+
+Agents have attributes. \textit{Name} is self-explanatory.
+\textit{Mechanization} is the agent's level of mechanization at a given
+timestep. Mechanization is calculated as the number of armored vehicles per 100
+soldiers, per \cite{sechser2010army}. \textit{Neighbors} is a list containing
+the \textit{Names} of an agent's geographic neighbors at a given point in
+time---more on this in the subsequent \textbf{Environment} section. \textit{NeighborsMech} is the
+average mechanization level of an agent's \textit{Neighbors} at a given point in
+time. \textit{Enemies} is a list containing the \textit{Names} of an agent's
+enemies at a given point in time. An enemy is an agent with which the agent in
+question has had a militarized interstate dispute in the preceding ten years,
+per the Correlates of War (COW) Militarized Interstate Disputes Database (MIDB)
+(\cite{cow_midb}). \textit{EnemiesMech} is the average mechanization score of
+enemies. Agents also have attributes that affect their
+perception of other agents and cognition for changing their own behavior. These
+attributes are discussed in the \textbf{Cognition and Actions} section.
 
@@
 Agents exist in a spatial environment. Spatiality is represented by neighbor
-relationships rather than a raster or vector GIS map. Two states are defined as
+relationships. Two states are defined as
 being neighbors if they share a contiguous land border or a stretch of water
-less than 400 miles. Every agent has an attribute \texttt{neighbor\_list}, a
-list in which it stores the names of neighboring agents for that time-step.
+less than 400 miles. Data on neighbor relationships is taken from the Correlates
+of War (COW) Direct Continuity dataset (\cite{cow_contiguity}).
+Modeling spatiality as dyadic neighbor relationships, rather than a more
+detailed raster or vector GIS implementation, is justified on the basis that the
+academic literature on mechanization suggests that two states neighboring each
+other is the important factor, not more detailed geographic information (e.g.,
+topographical features along the border). As such, this paper's modeling choice
+captures the phenomenon of interest while avoiding the overhead of a full-blown GIS implementation.
 
-\subsection{Initialization and Time}
+\subsection{Cognition}
 
-\subsection{Dynamics and Interactions}
+Agents have global perception of other agents: they can ``see'' the
+mechanization state of all other agents. This modeling choice is reflective of
+the real world, in which states devote significant resources to stay informed of
+the military posture of others. Agent cognition centers on 
+updating their own mechanization in response to the external security
+environment. This occurs in three ways. One, if an agent's
+\textit{NeighborsMech} increases---i.e., the agent's neighbors become more
+mechanized---the agent will increase its own mechanization by some proportional
+amount controlled by the agent's unique \textit{NeighborMechSensitivity}
+attribute, which is heterogenous across agents (more on this below).
+Two, if an agent's
+\textit{EnemiesMech} increases---i.e., the agent's enemies become more
+mechanized---the agent will increase its own mechanization by some proportional
+amount controlled by the agent's unique \textit{EnemyMechSensitivity} attribute,
+again heterogenous across agents. Three,
+agents occasionally suffer an ``anti-mechanization lesson'': defeat in a
+counter-insurgency conflict. Such events are scripted to occur for agents and
+years as per the historical record: e.g., the Soviet Union experienced such a
+lesson in the 1980s in Afghanistan. An agent experiencing such an event will
+decrease their own mechanization by some amount controlled by the agent's
+unique \textit{LessonSensitivity} attribute, again heterogenous across agents.
+
+An important note is that these three factors---neighbors, enemies, and
+defeat in counter-insurgency---correspond to the factors identified by 
+\cite{sechser2010army} as most important for mechanization. Because the research
+goal is essentially to see whether their findings stand up to agentization,
+it is critical to make agent cognition work in about the way that their findings
+predict. Inclusion of other factors not identified as important for
+mechanization---e.g., alliances, economic factors,
+culture, etc.---would be interesting but counter-productive to the research
+question. They are discussed along with this model's findings.
+
+Data for \textit{NeighborsMechSensitivity}, \textit{EnemiesMechSensitivity}, and
+\textit{LessonSensitivity} come from \cite{sechser2010army}. Specifically, they
+estimate coefficients and standard errors for these three factors, thus defining
+a statistical distribution for each. Upon instantiation, every agent makes a
+random draw from each of the three distributions, the resulting values of are
+taken for these three attributes. This introduces a stochastic element to the
+model and variation across runs, providing a richer body of data to draw upon.
+
+\subsection{Time}
+
+The model starts in 1979 and runs in 2-year time increments until 2001. This
+timespan and periodicity reflects data availability: a longer time period and
+more frequent updates would be ideal but is not practically achievable. Agent
+attributes are initialized, and then then the model iterates through timesteps.
+
+In discussing what happens each timestep, it is useful to distinguish between
+things that depend on agents' actions and those that do not.
+
+Some changes are deterministically programmed into the model and do not depend on
+agent behavior. Every timestep, neighbor relationships and enemy relationships
+are updated in accordance with the historical record. The occurrence of
+counter-insurgency defeats also is hardwired into the model and does not depend
+on agent behavior within the model.
+
+Other changes do depend on agent behavior. Agent activation order in a
+particular timestep is random. Every timestep, every agent perceives the world
+aorund them; recalculates \textit{NeighborsMech} and \textit{EnemiesMech};
+updates their mechanization level according to \textit{NeighborMechSensitivity}
+and \textit{EnemyMechSensitivity}; and updates their mechanization level
+according to \textit{LessonSensitivity} if an anti-mechanization lesson occured.
+
+\subsection{Output Data}
 

### 2018-05-11 paper2/prez.tex
o

\documentclass{beamer}
\usetheme{Madrid}
\usecolortheme{spruce}

\usepackage{graphicx}
\graphicspath{{./img/}}

\title{An agent-based model of military mechanization}
\subtitle{CSS 645}
\author{Tom Wallace}
\institute{George Mason University}
\date{Spring 2018}

\begin{document}

\frame{\titlepage}

\begin{frame}
	\frametitle{Military mechanization}

	States differ in the composition of their militaries, particularly ground forces \\~\\

	One axis of comparison is \textbf{mechanization}
	\begin{itemize}
		\item \small The degree to which an army is comprised of infantry vs. ground combat vehicles \\~\\
	\end{itemize}

	Historical examples help illustrate the general idea
	\begin{itemize}
		\item \small 1970s Vietcong (low mechanization) vs. 1980s Israel (high mechanization)
		\item \small 1980s U.S. (higher mechanization) vs. present-day U.S. (lower mechanization)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{The consequences of choices regarding force structure are great}
	Academic and professional consensus that choices regarding military
	composition (including but not limited to mechanization) condition
	battlefield effectiveness \\~\\

	Size, cost, and complexity dictate that states can only slowly alter 
	military composition, even under great duress \\~\\

	``You go to war with the army you have'' \\~\\
\end{frame}

\begin{frame}
	\frametitle{The determinants of mechanization are disputed}
	One view stresses \textbf{security environment}: states adjust
	mechanization to be able to prevail in conflict
	\begin{itemize}
		\item \small Neighbors' and enemys' mechanization (e.g. 1930s
			France and Germany)
		\item \small Terrain (e.g. Switzerland)
		\item \small Insurgency (e.g. U.S.)\\~\\
	\end{itemize}

	Another view stresses \textbf{political, economic, and cultural
	factors}: states are not pure security-maximizers
	\begin{itemize}
		\item \small Regime type (e.g. democracy vs. autocracy)
		\item \small Economic endowments (e.g. capital vs. labor)
		\item \small Cultural norms (e.g. Irish tanks)\\~\\
	\end{itemize}

	Sechser and Saunders 2010 conclude the former matters much more than the
	latter
	\begin{itemize}
		\item \small Developed state-year level dataset on
			mechanization, 1979-2001
		\item \small Traditional econometric analysis (OLS)
		\item \small Cultural norms (e.g. Irish tanks)\\~\\
	\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Viewing their findings through an ABM lens}

	Emphasis on \textbf{spatiality}: what are my neighbors doing? \\~\\

	My actions affect your actions affect my actions... $\to$ \textbf{positive feedback loops} \\~\\

	Estimated statistical coefficients implicitly claim particular cognitive model
	\begin{itemize}
		\item \small If neighbors mechanization goes up by $X$, I raise my mechanization by $\beta$ \\~\\
	\end{itemize}

	My research question: do their findings hold up under an ABM formulation?
\end{frame}

\begin{frame}
	\frametitle{Model description}
	Agent = state actor\\~\\

	Attributes:
	\begin{itemize}
		\item \small Mechanization level (vehicles : infantry ratio,
			IISS data)
		\item \small Geographic neighbors (contiguous border per COW dataset)
		\item \small Enemies (defined based on 10-year MID history)
		\item \small Terrain (Fearon and Laitin 2004) \\~\\
	\end{itemize}

	Cognition regarding mechanization
	\begin{itemize}
		\item \small Takes into account neighbors, enemies, terrain,
			wars
		\item \small Agents observe lagged global behavior (2 years prior) and calculate
			weighted average of neighbor mechanization; weighted
			average of enemy mechanization; own terrain; and wars
		\item \small For each of the above, adjust own mechanization by some
			coefficient
		\item \small Coefficients drawn from statistical distribution
			found by Sechser and Saunders, randomly assigned to
			agents (heterogeneity)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Model description}
	Initialized in 1979 \\~\\

	Proceeds in two-year timesteps until 2001 \\~\\

	Every turn, agents perceive neighbors, enemies, terrain, and wars, and
	adjust own mechanization accordingly \\~\\

	Because agent cognition has random element, model is stochastic and must
	be run many times \\~\\

	General research strategy: initialize with values from 1979, run, see if model results in 2001 match reality at both
	actor- and system-level
\end{frame}

\begin{frame}
	\frametitle{Example}
\end{frame}

\begin{frame}
	\frametitle{Findings}

	Model results in significantly higher mechanization at both system- and
	agent-level than reality

	INSERT IMAGES
\end{frame}

\begin{frame}
	\frametitle{Influence of positive feedback loops}

	Results stem from positive feedback loop: you increase your
	mechanization, which causes me to increase my mechanization, which
	causes you to increase your mechanization... \\~\\

	Modeled system lacks countervailing negative feedback loop \\~\\

	Conclusion: some of the factors dismissed by Sechser and Saunders must
	play a role in mitigating rate-of-increase in the real-world system
	\\~\\

	Cannot say \textit{which} factors, but the system-level point remains

\end{frame}

\begin{frame}
	\frametitle{Verification, validation, and other topics}

	In our taxonomy: analyzed-analyzed model \\~\\

	Validation central to research strategy: comparison of model-2001 to
	reality-2001 \\~\\

	Significant amount of coding work was dealing with fact that composition
	of international system changed over time (e.g., 1991)

	\begin{itemize}
		\item \small Inheritance
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Possible extensions}

	Alliances \\~\\

	More sophisticated cognition \\~\\

	Incorporation of other force structure choices (e.g. human capital) \\~\\
\end{frame}

\end{document}

### 2018-05-12 paper2/paper2.tex
foo

@@
 \documentclass{article}
+
+\usepackage{amsmath}
+\usepackage{graphicx}
+\graphicspath{}
 
@@
 reflexive process: states choose their mechanization posture largely based on
-that neighbors and enemies. This paper argues
+that neighbors, enemies, and recent military experiences (e.g. defeat in
+counter-insurgency). This paper argues
 that this view---based on traditional longitudinal data analysis---is ripe for
-agentization:
+agentization. It implicitly argues for autonomous agents (states) that have
+cognition regarding mechanization that takes into account other agents (e.g.
+enemies), spatiality (e.g. neighbors), and learning from experience. This paper
+builds such a model and validates it against the empirical record, using the
+National Mechanization Index of \cite{sechser2010army}. It finds that the 
+orthodox view of mechanization does not stand up to agentization. A positive feedback loop exists causes states to
+mechanize much faster and to a greater degree than occurs in the real-world.
+This implies the existence of some countervailing factor (possibly a negative
+feedback loop) that mitigates the extent of mechanization. Candidates for what
+this could be mostly center on ``internal'' factors---e.g., culture, economy,
+and so on---as opposed to the ``external'', security environment-centric
+orthodox view.
 
@@
 
-The consequences of choices regarding mechanization are large. There is
+The consequences of mechanization are large. There is
 academic consensus that defense policy---including but not limited to
@@
 leaders cannot always forecast conflict, and armies are large, complex
-organizations that typically are held to change only infrequently and slowly at
-that, even if there is a pressing need (\cite{murray1998military},
+organizations that typically are held to change only infrequently and slowly (\cite{murray1998military},
 \cite{locher2004victory}, \cite{zegart2000flawed}). A state suddenly faced with
@@
 that states are autonomous agents; that they have
-cognition regarding how to perceive the outside world in terms of mechanization
-and how to adjust their own mechanization to it; that spatiality and
-neighborhood matter; that inter-agent relationships (e.g. whether you are my
-enemy or not) matter; and that
+cognition regarding how to perceive the outside world
+and how to react to it; that spatiality and
+neighborhood matter; that inter-agent relationships matter; and that
 agents learn from experience. Agent-based modeling is uniquely well-suited to
@@
 I hypothesize that the arguments of \cite{sechser2010army} are incomplete and
-that some omitted variable or set of variables acts as a ``brake'' on military
+that some omitted variable(s) act(s) as a ``brake'' on 
 mechanization.\footnote{By ``omitted'', I mean that they conclude that the variable
@@
 states in the world at a particular point in time, minus microstates (defined as
-those with a population less than 750,000).
+those with a population less than 750,000). Some other states were excluded; for
+why, see the \textbf{Time}
+section. Figure 1 depicts the set of agents.
+
+\begin{figure}[h!]
+	\centering
+	\caption{Agents}
+\includegraphics[scale=0.05]{map.png}
+\end{figure}
 
@@
 topographical features along the border). As such, this paper's modeling choice
-captures the phenomenon of interest while avoiding the overhead of a full-blown GIS implementation.
+captures the phenomenon of interest while avoiding the overhead of a full-blown
+GIS implementation coupled with ABM
+(e.g., \cite{harper2002modeling}).
 
@@
 unique \textit{LessonSensitivity} attribute, again heterogenous across agents.
+Cognition for agent $i$ thus is:
+
+\begin{multline}
+\textrm{Mech}_i^{(t+1)} = \textrm{Mech}_i^{(t)} \\ 
++ (\textrm{NeighborMechSensitivity}_i \times \textrm{NeighborMech}_i^{(t)}) \\
++ (\textrm{EnemyMechSensitivity}_i \times \textrm{EnemyMech}_i^{(t)}) \\
+	+ (\textrm{LessonSensitivity}_i \times I_{\textrm{Lesson}}(t, i))
+\end{multline}
+
+with $(t)$ denoting timestep, $i$ indexing agents, and $I_{\textrm{Lesson}}(t,
+i)$ an indicator function of whether the agent experienced an anti-mechanization
+lesson. Table 1 summarizes agent attributes.
+
+\begin{table}[h]
+	\centering
+	\caption{Attributes, organized by theme}
+	\begin{tabular}{|l l|}
+		\hline
+		\textbf{Identity} & \textit{Name} \\
+		& \\
+		\textbf{Mechanization} & \textit{Mechanization} \\
+		& \\
+		\textbf{Spatiality} & \textit{Neighbors} \\
+		& \\
+		\textbf{Inter-agent Relationships} & \textit{Enemies} \\
+		& \\
+		\textbf{Perception} & \textit{EnemiesMech} \\
+		& \textit{NeighborsMech} \\
+		& \textit{Lesson} \\
+		& \\
+		\textbf{Cognition} & \textit{EnemyMechSensitivity} \\
+		& \textit{NeighborsMechSensitivity} \\
+		& \textit{LessonSensitivity} \\
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 random draw from each of the three distributions, the resulting values of are
-taken for these three attributes. This introduces a stochastic element to the
-model and variation across runs, providing a richer body of data to draw upon.
+taken for these three attributes. This modeling choice has two important
+elements. One, it further enhances agent heterogeneity. Agents already differ in
+their geographic neighbors, enemies, and experiences with anti-mechanization
+lessons; now, they also differ in the cognition used to update mechanization
+based on these factors. It also introduces a stochastic element to the
+model, meaning that each run differs and so multiple runs must be conducted for
+each specification.
 
@@
 attributes are initialized, and then then the model iterates through timesteps. 
-
 In discussing what happens each timestep, it is useful to distinguish between
@@
 
-\subsection{Output Data}
+An important caveat is that the set of agents represented in the model does not
+change over time: i.e., agents are the same from 1979 onwards. In real life,
+of course, the agent population was not static over this time period, with many
+states ceasing to exist in 1991 and new ones forming. Modeling these changes is
+beyond the scope of this model. How a collapsed state's military assets are divvied up
+among successor states is complex and not tractable to simple representation
+(e.g. averaging). For example, the Soviet Union deployed most of its modern
+military equipment in the far west of the country in anticipation of a conflict
+with NATO, and so when the USSR collapsed Ukraine inherited the lion's share, leaving
+little for successor states like Kazakhstan or Azerbaijan (\cite{roy1993military}). Such inheritance has
+a strong effect on the mechanization of of successor states. Designing
+complex and state-specific ``rules of inheritance'' was too much for this
+initial effort but is an area for further development.
+
+\subsection{Output Data, Verification, and Validation}
+
+The main model output of interest is data on military mechanization. It is provided for
+every agent and every time-step. These agent-level data also can be aggregated
+into system-level measures of military mechanization. Per the research design,
+the intent is to compare the model output to what actually occurred. If the
+model produces military mechanization levels much higher than reality, it means
+that Sechser and Saunder's findings do not stand up to agentization and there is
+some missing ``brake'' on military mechanization. Note that this research
+strategy has validation (comparison of model output to empirical data) at its
+core and so no separate validation is needed. Verification that the model is
+doing what it is intended was accomplished through manual inspection.
 
@@
 
+\subsection{System-Level}
+
+\subsection{Agent-Level}
+
 \section{Discussion}
+
+Learning ala \cite{bennett2006modelling}
+
 \printbibliography[heading=bibnumbered]

### 2018-05-13 paper2/paper2.tex
foo

@@
 
-\begin{centering}
-	Word count: XXXX
-\end{centering}
+\begin{center}
+	Word count: 4209
+\end{center}
 \section{Introduction}
@@
 GIS implementation coupled with ABM
-(e.g., \cite{harper2002modeling}).
+(e.g., \cite{harper2002modeling} or \cite{bennett2006modelling}).
 
@@
 with NATO, and so when the USSR collapsed Ukraine inherited the lion's share, leaving
-little for successor states like Kazakhstan or Azerbaijan (\cite{roy1993military}). Such inheritance has
+little for successor states such as Azerbaijan that were not important military
+districts (\cite{roy1993military}). Such inheritance has
 a strong effect on the mechanization of of successor states. Designing
@@
 
-\subsection{System-Level}
+At both the system- and agent-level, this agent-based model results in much
+higher mechanization levels than occurred in the real world. First, let us
+compare the average mechanization levels across all countries at each time-step
+of the model. Although the set of agents is not exactly the same between the
+model and reality (c.f. earlier issues with inheritance), this
+gross comparison still is a useful measure of comparison. Model results are
+the average of 100 runs, with all parameters (e.g.
+\textit{EnemyMechSensitivity}) drawn from the distributions claimed by
+\cite{sechser2010army}; empirical results are drawn from the National
+Mechanization Index compiled by \cite{sechser2010army}.
 
-\subsection{Agent-Level}
+\begin{table}[h]
+	\centering
+	\caption{Comparison}
+	\begin{tabular}{|l r r|}
+		\hline
+		\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
+		1979 & 0.0156 & 0.0139 \\
+		1981 & 0.0189 & 0.0167 \\
+		1983 & 0.0229 & 0.0167 \\
+		1985 & 0.0280 & 0.0183 \\
+		1987 & 0.0351 & 0.0187 \\
+		1989 & 0.0415 & 0.0197 \\
+		1991 & 0.0512 & 0.0215 \\
+		1993 & 0.0636 & 0.0229 \\
+		1995 & 0.0804 & 0.0231 \\
+		1997 & 0.1014 & 0.0270 \\
+		1999 & 0.1319 & 0.0272 \\
+		2001 & 0.1730 & 0.0291 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+\newpage
+
+Examining results at the agent-level reveals that the higher mechanization
+levels of the model are much more pronounced for some agents than others. In
+some cases, the model actually results in \textit{lower} mechanization levels,
+with Mexico serving as one example.
+
+\begin{table}[h]
+	\centering
+	\caption{Mexico}
+	\begin{tabular}{|l r r|}
+		\hline
+		\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
+		1979 & 0.0013 & 0.0012 \\
+		1981 & 0.0014 & 0.0012 \\
+		1983 & 0.0018 & 0.0026 \\
+		1985 & 0.0019 & 0.0013 \\
+		1987 & 0.0021 & 0.0019 \\
+		1989 & 0.0022 & 0.0027 \\
+		1991 & 0.0024 & 0.0030 \\
+		1993 & 0.0026 & 0.0033 \\
+		1995 & 0.0030 & 0.0035 \\
+		1997 & 0.0032 & 0.0074 \\
+		1999 & 0.0036 & 0.0082 \\
+		2001 & 0.0040 & 0.0075 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+For some agents, however, mechanization levels are extremely high. 
+Mechanization score is a percentage, and so values over 1 do not make sense, 
+yet some countries end the simulation with mechanization scores over 1. Note
+that one could easily right-bound values at 1, but this was not done in order to
+see what values would be organically converged to.
+
+\begin{table}[h]
+	\centering
+	\caption{Highest mechanization scores, 2001}
+	\begin{tabular}{|l r|}
+		\hline
+		\textbf{Agent} & \textbf{Mechanization Score} \\
+		Israel & 1.59 \\
+		USA & 1.34 \\
+		Kuwait & 1.14 \\
+		Russia & 1.06 \\
+		\hline
+	\end{tabular}
+\end{table}
+
+There appears to be a correlation between number of enemies and high
+mechanization score. For example, both the USA and Russia consistently had some
+of the highest amount of enemies, peaking with 16 enemies for Russia in 1999
+(this high number is attributable to the Kosovo crisis, when a confrontation between Russian and NATO forces
+at Pristina Airport nearly led to war), and as per the above table are among the
+most mechanized states. That Israel had many enemies during this time period is
+obvious.
+
+There also appears to be neighborhood effects. Mechanization is not evenly
+distributed across the globe. Rather, it is geographically concentrated in a few
+hot-spots. Agents belonging to the Middle East and North Africa (MENA) region
+account for about 10\% of the overall agent population, yet make up 50\% of the
+top 20 most mechanized agents at simulation end. Other regions
+are correspondingly under-represented; for example, Asia does not have a single
+state in the top 20.
+
+\begin{table}[h]
+	\centering
+	\caption{Top 20 mechanized agents at simulation end, by region}
+	\begin{tabular}{|l l l l l l|}
+		\hline
+		\textbf{MENA} & \textbf{Europe \& FSU} & \textbf{Africa} & \textbf{Americas} & \textbf{Asia} & \textbf{Oceania} \\
+		Israel       & Russia         & Sudan        & USA    & (none) & (none) \\
+		Kuwait       & France         & S. Africa    & Canada &        &        \\
+		Iran         & UK             &              &        &        &        \\
+		Libya        & Greece         &              &        &        &        \\
+		Iraq         & Netherlands    &              &        &        &        \\
+		Syria        & Poland         &              &        &        &        \\
+		KSA          &                &              &        &        &        \\
+		Egypt        &                &              &        &        &        \\
+		Lebanon      &                &              &        &        &        \\
+		Turkey       &                &              &        &        &        \\
+		\hline
+	\end{tabular}
+\end{table}
 
@@
 
-Learning ala \cite{bennett2006modelling}
+\subsection{Findings}
+
+Two conclusions should be drawn. One, the model results in much higher
+system-level mechanization than occurred in reality. Two, this higher
+mechanization is not evenly distributed across agents; rather, a small number of
+agents with many enemies and that all border each other become outliers and
+hence drive the global average far upward. The explanation for both phenomena
+lies in a positive feedback loop. Enemies and neighbors adjust their
+mechanization scores in response to each other; hence, a geographic cluster of
+enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
+mechanization.
+
+Three, because this model is a
+faithful representation of the factors that \cite{sechser2010army} claim are
+most important for mechanization, that model output fails to match reality
+indicates that their claimed model is incomplete. There must be some ``brake''
+on mechanization; put differently, some countervailing factor(s) must mitigate the
+effect of the positive feedback loop identified above. This model cannot
+conclusively identify what that ``brake'' is. The most plausible candidates are
+factors examined by \cite{sechser2010army} and deemed statistically
+insignificant such as regime type and economy, and factors not examined by
+\cite{sechser2010army} at all such as culture.
+
+This paper is a novel contribution to the literature. Although agent-based
+models have found traction in the study of grievance, insurgency, and
+other population-centric security phenomena (e.g., \cite{epstein2002modeling}),
+they quite rarely are used to model international relations and state-centric security
+phenomena despite some early work on this front (e.g., \cite{cederman1997emergent} and \cite{axelrod1997complexity}). 
+This paper's agentization of an orthodox security
+studies model thus is novel in method. Its findings also are important. They
+show that they show a solely ``external'', security
+environment-focused view of mechanization does not stand up to agentization.
+Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
+loops---all hallmarks of a complex system---drive results that significantly
+diverge from theoretical expectations of the orthodox view.
+
+\subsection{Areas for Improvement}
+
+Having taken a critical view of \cite{sechser2010army}, it only is fair that
+this paper applies that lens to itself. There are two main areas for
+improvement.
+
+One, this model's representation of \cite{sechser2010army} could be tightened
+up, as there currently is divergence between the two. For example, this model
+examines a smaller agent set than their paper. Building ``inheritance''
+processes to allow new states to emerge is desirable. This paper's
+use of statistical coefficients from \cite{sechser2010army} also has some
+methodological problems, a full description of which is omitted here due to
+space constraints. It suffices to say that a better approach would be to conduct
+more extensive parameter sweeps of relevant cognitive factors.
+
+Two, this model could incorporate potential explanations of what the missing
+``brake'' on mechanization might be. Economy and culture could be included in
+the model. Doing so would allow exploration of whether these factors are
+logically consistent and can act as a countervailing factor that controls the
+positive feedback loop of mechanization. It cannot \textit{prove} that these or
+any other factors actually exist in reality, but they can demonstrate how they
+are or are not consistent in a simulated reality.
+
+My hope is that making these changes will make this a publication-quality paper
+capable of serving as one of two required papers for the comprehensive exams.
+
+\newpage
 

### 2018-05-13 paper2/model.py

@@ -8,15 +8,9 @@ Final project
 import numpy as np
 import pandas as pd
 import math
+from random import shuffle
 
-def runSimulation():
-
-	coeff_enemies = 1.2
-	coeff_neighbors = 1.15
-	coeff_aml = 0.9
-	coeff_em = 1.04 #differs from their actual coefficient
-	coeff_nm = 1.02 #differs from their actual coefficient
-	coeff_mountains = 1.1
+def runSimulation(n_iter):
 
 	# import mechanization data
 	sechser = pd.read_csv("sechser-saunders-ISQ-2010.tab", sep = '\t')
@@ -42,36 +36,106 @@ def runSimulation():
 		agent = Agent(name)
 		lst.append(agent)
 
-	# get neighbors, enemies, aml, starting mech
-	for agent in lst:
-		agent.getNeighbors(1979, contdird)
-		agent.getEnemies(1979, rivals)
-		agent.getAML(1979, sechser)
-		agent.getStartingMech(1979, sechser)
+	# coeffs
+	coeff_enemy_mu = 0.209
+	coeff_enemy_ss = 0.081
 
-	# calculate enemy_mech and neighbor_mech
-	for agent in lst:
-		if(len(agent.enemies) > 0):
-			agent.calcEnemyMech(lst)
-			agent.calcNeighborMech(lst)
-	
-	# 1981
-	'''
-	for agent in lst:
-		print(agent.name, agent.current_mech, agent.mech, agent.neighbors, agent.neighbor_mech, agent.enemies,
-				agent.enemy_mech, agent.aml)
-	'''
+	coeff_enemy_mech_mu = 0.044
+	coeff_enemy_mech_ss = 0.019
+
+	coeff_neighbor_mu = 0.149
+	coeff_neighbor_ss = 0.059
+
+	coeff_neighbor_mech_mu = 0.023
+	coeff_neighbor_mech_ss = 0.013
+
+	coeff_aml_mu = -0.095
+	coeff_aml_ss = 0.044
+
+	for i in range(0, n_iter):
+
+		# get coeffs, starting mech
+		for agent in lst:
+			agent.drawNeighborSensitivity(coeff_neighbor_mu, coeff_neighbor_ss)
+			agent.drawNeighborMechSensitivity(coeff_neighbor_mech_mu,coeff_neighbor_mech_ss)
+
[…]

### 2018-05-14 paper2/paper2.tex
foo

@@
 \begin{center}
-	Word count: 4209
+	Word count: 4218
 \end{center}
@@
 National Mechanization Index of \cite{sechser2010army}. It finds that the 
-orthodox view of mechanization does not stand up to agentization. A positive feedback loop exists causes states to
+orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
+causes states to
 mechanize much faster and to a greater degree than occurs in the real-world.
@@
 of complex systems and again are well-suited for representation by agent-based
-modeling. This specific one---states changing their military in response to
-other states change in military in response to... and so on---feedback loop---is often called
+modeling. This specific feedback loop is often called
 the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
@@
 I hypothesize that the arguments of \cite{sechser2010army} are incomplete and
-that some omitted variable(s) act(s) as a ``brake'' on 
-mechanization.\footnote{By ``omitted'', I mean that they conclude that the variable
+that some omitted variable or set of variables acts as a ``brake'' on 
+mechanization.\footnote{By ``omitted'', I mean that Sechser and Saunders conclude that the variable
 is not very important for mechanization, not that they do not address the

### 2018-05-14 paper2/prez.tex
foo

@@
 \usetheme{Madrid}
-\usecolortheme{spruce}
 
@@
 \graphicspath{{./img/}}
+
+\AtBeginSection[]{
+	\begin{frame}
+		\frametitle{Table of Contents}
+		\tableofcontents[currentsection]
+	\end{frame}
+}
 
@@
 \frame{\titlepage}
+
+\section{Background}
 
@@
 			mechanization, 1979-2001
-		\item \small Traditional econometric analysis (OLS)
-		\item \small Cultural norms (e.g. Irish tanks)\\~\\
+		\item \small Traditional longitudinal data analysis (OLS)
 	\end{itemize}
@@
 
+\section{Model Description}
+
 \begin{frame}
-	\frametitle{Model description}
-	Agent = state actor\\~\\
+	\frametitle{Agents}
+	Agent = state actor \\~\\
 
-	Attributes:
+	Covers all countries in the world, minus microstates and ``inheritance'' problem
+
+	\begin{figure}[h!]
+		\centering
+		\caption{Agents}
+		\includegraphics[scale=0.04]{map.png}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Agent attributes}
+	\begin{table}[h]
+		\centering
+		\begin{tabular}{|l l|}
+			\hline
+			\textbf{Identity} & \textit{Name} \\
+			& \\
+			\textbf{Mechanization} & \textit{Mechanization} \\
+			& \\
+			\textbf{Spatiality} & \textit{Neighbors} \\
+			& \\
+			\textbf{Inter-agent Relationships} & \textit{Enemies} \\
+			& \\
+			\textbf{Perception} & \textit{EnemiesMech} \\
+			& \textit{NeighborsMech} \\
+			& \textit{Lesson} \\
+			& \\
+			\textbf{Cognition} & \textit{EnemyMechSensitivity} \\
+			& \textit{NeighborsMechSensitivity} \\
+			& \textit{LessonSensitivity} \\
+			\hline
+		\end{tabular}
+	\end{table}
+\end{frame}
+
+\begin{frame}
+	\frametitle{Additional notes on agents and environment}
+	Mechanization: combat vehicles per 100 infantry (data: IISS)\\~\\
+
+	Spatiality: neighbor relationships rather than full-blown GIS \\~\\
+
+	Enemies: defined on 10-year Militarized Interstate Dispute (MID) history\\~\\
+
+	Perception: every time step, agents observe behavior of others and respond\\~\\
+
+	Cognition: adjust own mechanization level based on neighbors, enemies, defeat in counter-insurgency
 	\begin{itemize}
-		\item \small Mechanization level (vehicles : infantry ratio,
-			IISS data)
-		\item \small Geographic neighbors (contiguous border per COW dataset)
-		\item \small Enemies (defined based on 10-year MID history)
-		\item \small Terrain (Fearon and Laitin 2004) \\~\\
-	\end{itemize}
-
-	Cognition regarding mechanization
-	\begin{itemize}
-		\item \small Takes into account neighbors, enemies, terrain,
-			wars
-		\item \small Agents observe lagged global behavior (2 years prior) and calculate
-			weighted average of neighbor mechanization; weighted
-			average of enemy mechanization; own terrain; and wars
-		\item \small For each of the above, adjust own mechanization by some
-			coefficient
-		\item \small Coefficients drawn from statistical distribution
-			found by Sechser and Saunders, randomly assigned to
-			agents (heterogeneity)
+		\item \small Cognition is heterogenous across agents:
+			coefficients drawn from distributions of Sechser and
+			Saunders 
+		\item \small Update equation:
+			\begin{multline}
+				\textrm{Mech}_i^{(t+1)} = \textrm{Mech}_i^{(t)}
+				+ (\textrm{NeighborMechSensitivity}_i \times \textrm{NeighborMech}_i^{(t)}) \\
+				+ (\textrm{EnemyMechSensitivity}_i \times \textrm{EnemyMech}_i^{(t)}) \\
+				+ (\textrm{LessonSensitivity}_i \times I_{\textrm{Lesson}}(t, i))
+			\end{multline}
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Model description}
-	Initialized in 1979 \\~\\
+	\frametitle{Time}
+	Initialized in 1979, proceeds in two-year timesteps until 2001 \\~\\
 
-	Proceeds in two-year timesteps until 2001 \\~\\
-
-	Every turn, agents perceive neighbors, enemies, terrain, and wars, and
-	adjust own mechanization accordingly \\~\\
+	Every turn, agents (in random activation order):
+	\begin{itemize}
+		\item \small Perceive world
+		\item \small Update assessment of neighbors, enemies, AML
+		\item \small Adjust own mechanization \\~\\
+	\end{itemize}
 
@@
 
+\section{Findings}
+
 \begin{frame}
-	\frametitle{Example}
+	\frametitle{Model mechanization much higher at system-level than
+	occurred in reality}
+\begin{table}[h]
+	\centering
+	\caption{Comparison}
+	\begin{tabular}{|l r r|}
+		\hline
+		\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
+		1979 & 0.0156 & 0.0139 \\
+		1981 & 0.0189 & 0.0167 \\
+		1983 & 0.0229 & 0.0167 \\
+		1985 & 0.0280 & 0.0183 \\
+		1987 & 0.0351 & 0.0187 \\
+		1989 & 0.0415 & 0.0197 \\
+		1991 & 0.0512 & 0.0215 \\
+		1993 & 0.0636 & 0.0229 \\
+		1995 & 0.0804 & 0.0231 \\
+		1997 & 0.1014 & 0.0270 \\
+		1999 & 0.1319 & 0.0272 \\
+		2001 & 0.1730 & 0.0291 \\
+		\hline
+	\end{tabular}
+\end{table}
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Findings}
-
-	Model results in significantly higher mechanization at both system- and
-	agent-level than reality
-
-	INSERT IMAGES
+	\frametitle{Mechanization not uniform across agents - actually
+	\textit{lower} in some}
+	\begin{table}[h]
+		\centering
+		\caption{Mexico}
+		\begin{tabular}{|l r r|}
+			\hline
+			\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
+			1979 & 0.0013 & 0.0012 \\
+			1981 & 0.0014 & 0.0012 \\
+			1983 & 0.0018 & 0.0026 \\
+			1985 & 0.0019 & 0.0013 \\
+			1987 & 0.0021 & 0.0019 \\
+			1989 & 0.0022 & 0.0027 \\
+			1991 & 0.0024 & 0.0030 \\
+			1993 & 0.0026 & 0.0033 \\
+			1995 & 0.0030 & 0.0035 \\
+			1997 & 0.0032 & 0.0074 \\
+			1999 & 0.0036 & 0.0082 \\
+			2001 & 0.0040 & 0.0075 \\
+			\hline
+		\end{tabular}
+	\end{table}
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Influence of positive feedback loops}
-
-	Results stem from positive feedback loop: you increase your
-	mechanization, which causes me to increase my mechanization, which
-	causes you to increase your mechanization... \\~\\
-
-	Modeled system lacks countervailing negative feedback loop \\~\\
-
-	Conclusion: some of the factors dismissed by Sechser and Saunders must
-	play a role in mitigating rate-of-increase in the real-world system
-	\\~\\
-
-	Cannot say \textit{which} factors, but the system-level point remains
-
+	\frametitle{Findings driven by few outliers}
+	\begin{table}[h]
+		\centering
+		\caption{Highest mechanization scores, 2001}
+		\begin{tabular}{|l r|}
+			\hline
+			\textbf{Agent} & \textbf{Mechanization Score} \\
+			Israel & 1.59 \\
+			USA & 1.34 \\
+			Kuwait & 1.14 \\
+			Russia & 1.06 \\
+			\hline
+		\end{tabular}
+	\end{table}
 \end{frame}
@@
 \begin{frame}
-	\frametitle{Verification, validation, and other topics}
+	\frametitle{Outliers are clustered by neighborhood and relationship}
+	\begin{table}[h]
+		\centering
+		\caption{Top 20 mechanized agents at simulation end, by region}
+		\begin{tabular}{|l l l l l l|}
+			\hline
+			\textbf{MENA} & \textbf{Europe \& FSU} & \textbf{Africa} & \textbf{Americas} & \textbf{Asia} & \textbf{Oceania} \\
+			Israel       & Russia         & Sudan        & USA    & (none) & (none) \\
+			Kuwait       & France         & S. Africa    & Canada &        &        \\
+			Iran         & UK             &              &        &        &        \\
+			Libya        & Greece         &              &        &        &        \\
+			Iraq         & Netherlands    &              &        &        &        \\
+			Syria        & Poland         &              &        &        &        \\
+			KSA          &                &              &        &        &        \\
+			Egypt        &                &              &        &        &        \\
+			Lebanon      &                &              &        &        &        \\
+			Turkey       &                &              &        &        &        \\
+			\hline
+		\end{tabular}
+	\end{table}
+\end{frame}
 
-	In our taxonomy: analyzed-analyzed model \\~\\
+\section{Discussion}
 
-	Validation central to research strategy: comparison of model-2001 to
-	reality-2001 \\~\\
+\begin{frame}
+	\frametitle{Findings driven by positive feedback loop}
 
-	Significant amount of coding work was dealing with fact that composition
-	of international system changed over time (e.g., 1991)
+	State A raises mechanization, which causes neighbor / enemy State B to
+	increase own mechanization... (\textit{security dilemma})\\~\\
 
+	States with high exposure to this feedback loop due to many enemies and
+	neighbors become outliers \\~\\
+
+	Positive feedback loops are hallmark of complex system
 	\begin{itemize}
-		\item \small Inheritance
+		\item \small Autonomous, cognitive, heterogenous agents
+		\item \small Spatiality
+		\item \small Interaction
 	\end{itemize}
@@
 \begin{frame}
-	\frametitle{Possible extensions}
+	\frametitle{Implications for larger literature}
+	Findings imply orthodox, security environment-centric, ``external'' view
+	is incomplete: there must be some variable or set of variables acting as
+	a ``brake'' on military mechanization to mitigate the aforementioned
+	positive feedback loop \\~\\
 
-	Alliances \\~\\
+	Model cannot say exactly what those missing variables (but could assess
+	whether candidates ``work'' in simulation...hint hint) \\~\\
 
-	More sophisticated cognition \\~\\
+	Agent-based models relatively rare in international relations (e.g.
+	Cederman 1997, Axelrod 1997)
+\end{frame}
+\begin{frame}
+	\frametitle{Larger course themes}
 
-	Incorporation of other force structure choices (e.g. human capital) \\~\\
+	Analyzed-analyzed\\~\\
+
+	Empirical methods and validation central to research strategy \\~\\
+
+	GIS in ABM: what level of detail is good enough? \\~\\
+
+	Understanding behavior of complex models \\~\\
+
 \end{frame}

### 2018-07-08 paper2/paper2.tex
update

@@
 
-\usepackage[citestyle=authoryear, style=authoryear]{biblatex}
+\usepackage[citestyle=authoryear,style=authoryear,backend=biber,natbib]{biblatex}
 \addbibresource{paper2.bib}
@@
 
-\begin{center}
-	Word count: 4218
-\end{center}
+\begin{abstract}
+
+Military mechanization is the degree to which a state's army is comprised of
+ground combat vehicles as opposed to infantry. Mechanization is both a 
+theoretically important topic in international relations scholarship and a practical 
+concern for policymakers. \citet{sechser2010army} influentially argue that
+inter-state dynamics (e.g., rivalries) better explain mechanization choices than do
+intra-state factors (e.g., culture). This paper demonstrates that
+the multivariate statistical approach employed in \citet{sechser2010army} is inadequate. 
+It presents an alternate agent-based model of military mechanization that has superior explanatory power.
+
+\end{abstract}
+
+\newpage
+
 \section{Introduction}
@@
 builds such a model and validates it against the empirical record, using the
-National Mechanization Index of \cite{sechser2010army}. It finds that the 
+National Mechanization Index of \citet{sechser2010army}. It finds that the 
 orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
@@
 mechanization---conditions battlefield effectiveness. 
-\cite{lyall2009rage} finds that highly mechanized militaries are less effective
-at counterinsurgency, while \cite{biddle2004military} argues that
+\citet{lyall2009rage} finds that highly mechanized militaries are less effective
+at counterinsurgency, while \citet{biddle2004military} argues that
 mechanization-enabled mobility aids conventional warfighting. It thus
@@
 leaders cannot always forecast conflict, and armies are large, complex
-organizations that typically are held to change only infrequently and slowly (\cite{murray1998military},
-\cite{locher2004victory}, \cite{zegart2000flawed}). A state suddenly faced with
+organizations that typically are held to change only infrequently and slowly
+\citep{murray1998military,locher2004victory,zegart2000flawed}. A state suddenly faced with
 a war for which its mechanization posture is ill-suited can either concede or
@@
 mechanization---emphasizes the point:
-``you go to war with the army you have'' (\cite{schmitt_2004}).
+``you go to war with the army you have'' \citep{schmitt_2004}.
 
@@
 signal of deterrence, or to learn from the perceived mistakes of the last war
-(\cite{mearsheimer1983conventional};
-\cite{huth1988extended}; \cite{murray2011military}). The other school of thought is more
+\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. The other school of thought is more
 diverse and so harder to name, but may be termed 
@@
 coup-prone, the state of civil-military relations, and so on---may influence 
-force structure (\cite{reiter2002democracies}; \cite{quinlivan1999coup};
-\cite{talmadge2015dictator}; \cite{brooks2008shaping}). Economic factor
+force structure \citep{reiter2002democracies,quinlivan1999coup,talmadge2015dictator,brooks2008shaping}. 
+Economic factor
 endowments may play a role: e.g., capital-rich states may mechanize more
-than labor-rich states (\cite{gartzke2001democracy}). Lastly,
-ideology (\cite{van1984cult}) and culture (e.g., \cite{pollack2004arabs}) are
+than labor-rich states \citep{gartzke2001democracy}. Lastly,
+ideology \citep{van1984cult} and culture \citep[e.g.][]{pollack2004arabs} are
 hard-to-measure but perhaps influential. Consider Ireland: between the 1920s and
@@
 professional military, an important factor for the young Irish state
-(\cite{farrell1998professionalization};
-\cite{farrell2001transnational}).
+\citep{farrell1998professionalization,farrell2001transnational}.
 
-\cite{sechser2010army} is the landmark study on military mechanization. They
+\citet{sechser2010army} is the landmark study on military mechanization. They
 assemble a longitudinal dataset at the country-year level of the mechanization
@@
 
-The findings of \cite{sechser2010army} are ripe for agentization. The
+The findings of \citet{sechser2010army} are ripe for agentization. The
 conclusions drawn implicitly claim
@@
 agents learn from experience. Agent-based modeling is uniquely well-suited to
-represent such phenomena (\cite{gilbert2005simulation};
-\cite{miller2009complex}).
+represent such phenomena \citep{gilbert2005simulation,miller2009complex}.
 
-Importantly, the findings of \cite{sechser2010army} imply a positive feedback
+Importantly, the findings of \citet{sechser2010army} imply a positive feedback
 loop. State A has high
@@
 the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
-of agent-based modeling (\cite{schelling1960strategy};
-\cite{schelling2006micromotives}).
+of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.
 
@@
 
-I hypothesize that the arguments of \cite{sechser2010army} are incomplete and
+I hypothesize that the arguments of \citet{sechser2010army} are incomplete and
 that some omitted variable or set of variables acts as a ``brake'' on 
@@
 terms, agents (states) are defined to have the cognitive and learning behavior
-implicitly claimed by \cite{sechser2010army}. The model is initialized with
+implicitly claimed by \citet{sechser2010army}. The model is initialized with
 real-world data from 1979 and allowed to run until 2001. The micro- and
 macro-outcomes are then validated against real-world 2001 data. If
-\cite{sechser2010army} are correct, the model's 2001 outcomes should match
+\citet{sechser2010army} are correct, the model's 2001 outcomes should match
 real-world 2001 outcomes. If the model mechanization levels in 2001 are much
@@
 This section of the paper describes the model. It does not use the ODD protocol
-of \cite{grimm2006standard} but covers many of the same points.
+of \citet{grimm2006standard} but covers many of the same points.
 
@@
 timestep. Mechanization is calculated as the number of armored vehicles per 100
-soldiers, per \cite{sechser2010army}. \textit{Neighbors} is a list containing
+soldiers, per \citet{sechser2010army}. \textit{Neighbors} is a list containing
 the \textit{Names} of an agent's geographic neighbors at a given point in
@@
 per the Correlates of War (COW) Militarized Interstate Disputes Database (MIDB)
-(\cite{cow_midb}). \textit{EnemiesMech} is the average mechanization score of
+\citep{cow_midb}. \textit{EnemiesMech} is the average mechanization score of
 enemies. Agents also have attributes that affect their
@@
 less than 400 miles. Data on neighbor relationships is taken from the Correlates
-of War (COW) Direct Continuity dataset (\cite{cow_contiguity}).
+of War (COW) Direct Continuity dataset \citep{cow_contiguity}.
 Modeling spatiality as dyadic neighbor relationships, rather than a more
@@
 captures the phenomenon of interest while avoiding the overhead of a full-blown
-GIS implementation coupled with ABM
-(e.g., \cite{harper2002modeling} or \cite{bennett2006modelling}).
+GIS implementation coupled with ABM \citep[e.g.][]{harper2002modeling,bennett2006modelling}.
 
@@
 defeat in counter-insurgency---correspond to the factors identified by 
-\cite{sechser2010army} as most important for mechanization. Because the research
+\citet{sechser2010army} as most important for mechanization. Because the research
 goal is essentially to see whether their findings stand up to agentization,
@@
 Data for \textit{NeighborsMechSensitivity}, \textit{EnemiesMechSensitivity}, and
-\textit{LessonSensitivity} come from \cite{sechser2010army}. Specifically, they
+\textit{LessonSensitivity} come from \citet{sechser2010army}. Specifically, they
 estimate coefficients and standard errors for these three factors, thus defining
@@
 little for successor states such as Azerbaijan that were not important military
-districts (\cite{roy1993military}). Such inheritance has
+districts \citep{roy1993military}. Such inheritance has
 a strong effect on the mechanization of of successor states. Designing
@@
 \textit{EnemyMechSensitivity}) drawn from the distributions claimed by
-\cite{sechser2010army}; empirical results are drawn from the National
-Mechanization Index compiled by \cite{sechser2010army}.
+\citet{sechser2010army}; empirical results are drawn from the National
+Mechanization Index compiled by \citet{sechser2010army}.
 
@@
 Three, because this model is a
-faithful representation of the factors that \cite{sechser2010army} claim are
+faithful representation of the factors that \citet{sechser2010army} claim are
 most important for mechanization, that model output fails to match reality
@@
 conclusively identify what that ``brake'' is. The most plausible candidates are
-factors examined by \cite{sechser2010army} and deemed statistically
+factors examined by \citet{sechser2010army} and deemed statistically
 insignificant such as regime type and economy, and factors not examined by
-\cite{sechser2010army} at all such as culture.
+\citet{sechser2010army} at all such as culture.
 
@@
 models have found traction in the study of grievance, insurgency, and
-other population-centric security phenomena (e.g., \cite{epstein2002modeling}),
+other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
 they quite rarely are used to model international relations and state-centric security
-phenomena despite some early work on this front (e.g., \cite{cederman1997emergent} and \cite{axelrod1997complexity}). 
+phenomena despite some early work on this front
+\citep[see][]{cederman1997emergent,axelrod1997complexity}. 
 This paper's agentization of an orthodox security
@@
 
-Having taken a critical view of \cite{sechser2010army}, it only is fair that
+Having taken a critical view of \citet{sechser2010army}, it only is fair that
 this paper applies that lens to itself. There are two main areas for
@@
 
-One, this model's representation of \cite{sechser2010army} could be tightened
+One, this model's representation of \citet{sechser2010army} could be tightened
 up, as there currently is divergence between the two. For example, this model
@@
 processes to allow new states to emerge is desirable. This paper's
-use of statistical coefficients from \cite{sechser2010army} also has some
+use of statistical coefficients from \citet{sechser2010army} also has some
 methodological problems, a full description of which is omitted here due to

### 2018-07-08 paper2/paper2.tex
sources

@@
 \newpage
+
+Methodology studies
+\citep{pepinsky2005agents,de2014agent,gunitsky2013complexity,bousquet2011beyond,harrison2012complexity,jervis1998system,modelski1990world,modelski1999long}
+
+Insurgency studies
+\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
+
+Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
+
+Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
+
+Just because \citep{epstein1996growing}
+
+Maybe it's because states are not unitary. Not necessariy requirement: e.g.,
+household is agent in Schelling model.
 

### 2018-07-08 paper2/paper2.tex
thinking

@@
 Methodology studies
-\citep{pepinsky2005agents,de2014agent,gunitsky2013complexity,bousquet2011beyond,harrison2012complexity,jervis1998system,modelski1990world,modelski1999long}
+\citep{pepinsky2005agents,de2014agent,
+gunitsky2013complexity,bousquet2011beyond,
+harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
 
@@
 household is agent in Schelling model.
+
+Cite Cederman to say why I don't do inheritenace - reification of actor
+
+Why ABM:
+ - inadequacy of SS - VIF, linear model
+ - uncertainty of causation
+ - endogeneity
+ - GDP / capita explains this...but also everything else
+ - have to go with simulation
+ - bigger picture, world politics as complex system
+ - look at Axtell \citep{axtell2000agents}
 
@@
 	\caption{Comparison}
-	\begin{tabular}{|l r r|}
+	\begin{tabular}{|l r r r|}
 		\hline
-		\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
-		1979 & 0.0156 & 0.0139 \\
-		1981 & 0.0189 & 0.0167 \\
-		1983 & 0.0229 & 0.0167 \\
-		1985 & 0.0280 & 0.0183 \\
-		1987 & 0.0351 & 0.0187 \\
-		1989 & 0.0415 & 0.0197 \\
-		1991 & 0.0512 & 0.0215 \\
-		1993 & 0.0636 & 0.0229 \\
-		1995 & 0.0804 & 0.0231 \\
-		1997 & 0.1014 & 0.0270 \\
-		1999 & 0.1319 & 0.0272 \\
-		2001 & 0.1730 & 0.0291 \\
+		\textbf{Year} & \textbf{Model} & \textbf{Reality} & \textbf{Difference}\\
+		1979 & 0.0156 & 0.0139 & 0.0017 \\
+		1981 & 0.0189 & 0.0167 & 0.0022 \\
+		1983 & 0.0229 & 0.0167 & 0.0062 \\
+		1985 & 0.0280 & 0.0183 & 0.0097 \\
+		1987 & 0.0351 & 0.0187 & 0.0164 \\
+		1989 & 0.0415 & 0.0197 & 0.0218 \\
+		1991 & 0.0512 & 0.0215 & 0.0297 \\
+		1993 & 0.0636 & 0.0229 & 0.0407 \\
+		1995 & 0.0804 & 0.0231 & 0.0573 \\
+		1997 & 0.1014 & 0.0270 & 0.0744 \\
+		1999 & 0.1319 & 0.0272 & 0.1047 \\
+		2001 & 0.1730 & 0.0291 & 0.1439 \\
 		\hline
@@
 	\caption{Mexico}
-	\begin{tabular}{|l r r|}
+	\begin{tabular}{|l r r r|}
 		\hline
-		\textbf{Year} & \textbf{Model} & \textbf{Reality} \\
-		1979 & 0.0013 & 0.0012 \\
-		1981 & 0.0014 & 0.0012 \\
-		1983 & 0.0018 & 0.0026 \\
-		1985 & 0.0019 & 0.0013 \\
-		1987 & 0.0021 & 0.0019 \\
-		1989 & 0.0022 & 0.0027 \\
-		1991 & 0.0024 & 0.0030 \\
-		1993 & 0.0026 & 0.0033 \\
-		1995 & 0.0030 & 0.0035 \\
-		1997 & 0.0032 & 0.0074 \\
-		1999 & 0.0036 & 0.0082 \\
-		2001 & 0.0040 & 0.0075 \\
+		\textbf{Year} & \textbf{Model} & \textbf{Reality} & \textbf{Difference}\\
+		1979 & 0.0013 & 0.0012 & 0.0001  \\
+		1981 & 0.0014 & 0.0012 & 0.0002  \\
+		1983 & 0.0018 & 0.0026 & -0.0008 \\
+		1985 & 0.0019 & 0.0013 & 0.0006  \\
+		1987 & 0.0021 & 0.0019 & 0.0002  \\
+		1989 & 0.0022 & 0.0027 & -0.0005 \\
+		1991 & 0.0024 & 0.0030 & -0.0006 \\
+		1993 & 0.0026 & 0.0033 & -0.0007 \\
+		1995 & 0.0030 & 0.0035 & -0.0005 \\
+		1997 & 0.0032 & 0.0074 & -0.0042 \\
+		1999 & 0.0036 & 0.0082 & -0.0046 \\
+		2001 & 0.0040 & 0.0075 & -0.0035 \\
 		\hline

### 2018-07-10 paper2/paper2.tex
multicollinearity

@@
 intra-state factors (e.g., culture). This paper demonstrates that
-the multivariate statistical approach employed in \citet{sechser2010army} is inadequate. 
-It presents an alternate agent-based model of military mechanization that has superior explanatory power.
+the multivariate statistical approach employed in \citet{sechser2010army} is
+	inadequate and that agent-based modeling is needed to capture
+	international security competition as a complex system. Candidate
+	agent-based models are presented.
 
@@
 
+Two central claims of this paper are that the multivariate statistical approach
+of \citet{sechser2010army} is inadequate and agent-based modeling is needed.
+They are addressed in turn.
+
+\citet{sechser2010army} gather observational panel data that consists of a measure 
+of military mechanization $y$ indexed by state $i$ and observation time $k$, as well as
+time-indexed covariates $x$. The association of covariates with mechanization is
+estimated via ordinary least squares (OLS) linear regression:
+
+\begin{equation}
+	y_{ik} = \beta_0 + \beta_j x_{ijk} + \epsilon \hspace{1em}
+\end{equation}
+
+Multicollinearity - everything correlated with each other
+
+The fitted model attempts to characterize the separate
+effect of each factor upon mechanization. However, evidence suggests that
+the covariates in fact are strongly cross-correlated. The problem of
+multi-collinearity can be expressed both
+
 \section{Introduction}
+
+\subsection{Outline}
 
@@
 high mechanization.
+
+Practical and academic consequences
 

### 2018-07-12 paper2/paper2.tex
background

@@
 	concern for policymakers. \citet{sechser2010army} influentially argue that
-inter-state dynamics (e.g., rivalries) better explain mechanization choices than do
+	inter-state dynamics (e.g., conflict) better explain mechanization choices than do
 	intra-state factors (e.g., culture). This paper demonstrates that
@@
 
+Things they don't capture: selection into war and deterrence, gdp / capita to
+political structure, institutional culture and choice, network effects and
+emulation
+
 Two central claims of this paper are that the multivariate statistical approach
@@
 
-Multicollinearity - everything correlated with each other
+Multicollinearity - everything correlated with each other.
 
@@
 
-This paper presents an agent-based model of military mechanization. First, it
-provides a theoretical discussion of military mechanization, states the research
-hypothesis, and explains why agent-based modeling is an appropriate research strategy.
-Then, it details an original agent-based model; presents results; and discusses
-their implications for the academic debate on military mechanization.
+This paper presents an agent-based model of military mechanization. It
+reviews the academic literature on military mechanization. It discusses why
+orthodox modeling strategies fail to capture international security competition
+as a complex system and why agent-based modeling is needed. In support of these
+claims, it ``agentizes'' the leading model of military mechanization and shows that the
+results of doing cast doubt on it. Finally, it presents a
+novel agent-based model that better represents military mechanization.
 
-In summary, the orthodox view of military mechanization is that it is a
+In summary, the orthodox view of mechanization is that it is a
 reflexive process: states choose their mechanization posture largely based on
@@
 
-\section{Background}
+\subsection{Background}
 
-\subsection{Military mechanization}
-
-States differ in the composition of their military ground forces.\footnote{States may have multiple organizations that
+States differ in the composition of their ground forces.\footnote{States may have multiple organizations that
 conduct ground combat operations: e.g., the United States operates both an Army
@@
 one extreme is an army consisting exclusively of dismounted, small 
-arms-wielding infantry units; the other extreme is an army consisting 
-solely of tank units, armored personnel carriers, artillery, and the like.
-Clearly, no modern state embodies either extreme, but the 1970s-era Vietcong are
+arms-wielding infantry units. The other extreme is an army consisting 
+solely of tanks, armored personnel carriers, and the like.
+No modern state fully embodies either extreme, but the 1970s-era Vietcong are
 a good example of a low-mechanization army and the 1980s-era Israeli army of
@@
 
-Practical and academic consequences
+Mechanization is important for multiple academic literatures. A
+fundamental debate in international relations (IR) is whether state
+behavior is driven by structural insecurity and 
+competition \citep{waltz2001man, mearsheimer2001tragedy}; or, whether other
+factors such as domestic political structure \citep{russett1994grasping} and
+socially constructed beliefs \citep{wendt1992anarchy} can point states to more peaceful directions. 
+The extent to which patterns of mechanization do or do not correlate with
+\textit{realpolitik} preparation for war informs this debate.
+The security studies community is similarly split on the question of military
+change. Some view militaries (particularly large ones) as change-averse, 
+inertia-driven bureaucracies that can only be bestirred by major events (e.g.,
+war or political reform)
+\citep{posen1986sources,zegart2000flawed,locher2004victory,halperin2007bureaucratic}. 
+Others contend that militaries can and do radically self-transform in both composition and doctrine
+\citep{murray1998military,adamsky2010culture}. The patterns of
+mechanization---stable? punctuated by sudden change? constantly evolving?---thus
+have strong implications for this debate.
 
-The consequences of mechanization are large. There is
-academic consensus that defense policy---including but not limited to
-mechanization---conditions battlefield effectiveness. 
-\citet{lyall2009rage} finds that highly mechanized militaries are less effective
+Mechanization choices also have more visceral impact.
+\citet{lyall2009rage} find that highly mechanized militaries are less effective
 at counterinsurgency, while \citet{biddle2004military} argues that
 mechanization-enabled mobility aids conventional warfighting. It thus
-pays to have an army that is designed to win the
-kinds of
-conflicts that a state is likely to face. Lacking a crystal ball, however,
-leaders cannot always forecast conflict, and armies are large, complex
-organizations that typically are held to change only infrequently and slowly
-\citep{murray1998military,locher2004victory,zegart2000flawed}. A state suddenly faced with
-a war for which its mechanization posture is ill-suited can either concede or
-pay a cost in blood and treasure; what it cannot do is
-quickly overcome the constraints imposed by past mechanization decisions. 
-A quote from then-Secretary of Defense Donald Rumsfeld---made in 2004, when the U.S. military
-was confronted in Iraq with the consequences of past choices regarding
-mechanization---emphasizes the point:
-``you go to war with the army you have'' \citep{schmitt_2004}.
+pays to have an army that is designed to win the kinds of conflicts that a state is 
+likely to face. Incorrect forecasting and choosing the wrong mechanization
+posture costs blood and treasure. The experience of the U.S. military in the early
+2000s, when it shifted under duress from a highly mechanized Cold War-legacy force to a less
+mechanized counter-insurgency one, provides a vivid example. A quote from then-Secretary of Defense 
+Donald Rumsfeld emphasizes the point:
+``you go to war with the army you have, not the army you might wish to have'' \citep{schmitt_2004}.
 
-The causes of military mechanization are disputed. \textit{Realists} believe 
-that mechanization choices are driven by security environment. States
-choose mechanization policies that they believe will allow them to prevail against potential adversaries, or to send a
+There are many candidates for the causes of mechanization. Inter-state
+competition and war is one. States may choose choose mechanization policies 
+that they believe will allow them to prevail against potential adversaries, or to send a
 signal of deterrence, or to learn from the perceived mistakes of the last war
-\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. The other school of thought is more
-diverse and so harder to name, but may be termed 
-\textit{institutionalists}. They believe that factors other than 
-strategic calculation determine military mechanization. Domestic political
-institutions---e.g., democracy vs. autocracy, politically stable vs.
-coup-prone, the state of civil-military relations, and so on---may influence 
-force structure \citep{reiter2002democracies,quinlivan1999coup,talmadge2015dictator,brooks2008shaping}. 
-Economic factor
-endowments may play a role: e.g., capital-rich states may mechanize more
-than labor-rich states \citep{gartzke2001democracy}. Lastly,
-ideology \citep{van1984cult} and culture \citep[e.g.][]{pollack2004arabs} are
-hard-to-measure but perhaps influential. Consider Ireland: between the 1920s and
-1940s, it expended a large portion of its defense budget on building
-a small number of tanks. They were wholly insufficient to repel a British 
-invasion (their ostensible purpose), and
-Ireland would have been better-served to pursue a low-mechanization guerrilla
-defense strategy, but tanks were seen as a prestigious signifier of a ``real''
-professional military, an important factor for the young Irish state
+\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. 
+Economic endowments may play a role: e.g., capital-rich states may mechanize more
+than labor-rich states \citep{gartzke2001democracy}. Domestic political
+institutions may matter: e.g., autocracies may be more willing than democracies to adopt
+low-mechanization ``human wave'' tactics
+\citep{quinlivan1999coup,reiter2002democracies,brooks2008shaping,talmadge2015dictator}. 
+Ideology \citep{van1984cult} and national culture \citep[e.g.][]{pollack2004arabs} are
+hard-to-measure but perhaps influential. Lastly,
+organizational culture can play a role \citep{builder1989masks}. Militaries may pursue a
+certain mechanization posture because it aligns with their institutional
+objectives, which may or may not be coincident with rational strategy. 
+Consider Ireland: between the 1920s and 1940s, it seriously considered expending a 
+large portion of its defense budget on building a relatively small number of tanks. 
+They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
+and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
+but tanks were seen as a prestigious signifier of a ``real'' professional
+military and hence desirable
 \citep{farrell1998professionalization,farrell2001transnational}.
@@
 assemble a longitudinal dataset at the country-year level of the mechanization
-of all states between 1979 and 2001, and conduct regression analysis to identify
-the effect of covariates on mechanization. They conclude that ``choices about
+of all states between 1979 and 2001. They conduct regression analysis to identify
+the effect of covariates on mechanization. They conclude that out of the
+aforementioned potential causes---security environment, economy, domestic
+politics, and ideology / culture---that the first is most important: ``choices about
 mechanization are strongly associated with a state's security environment.''
-The more mechanized a state's geographic neighbors and enemies, the higher the state's 
-own mechanization; also, states
-learn from defeat in insurgency by subsequently decreasing their mechanization.
 
@@
 
-The findings of \citet{sechser2010army} are ripe for agentization. The
+There is reason to believe that the orthodox multivariate statistical modeling 
+approach employed in \citet{sechser2010army} is insufficient. Several factors point towards 
+agent-based modeling being a more appropriate choice.
+
+One, from a stat
+
+are ripe for agentization. The
 conclusions drawn implicitly claim

### 2018-07-12 paper2/paper2.tex
progress

@@
 
-Methodology studies
-\citep{pepinsky2005agents,de2014agent,
-gunitsky2013complexity,bousquet2011beyond,
-harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
-
-Insurgency studies
-\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
-
-Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
-
-Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
-
-Just because \citep{epstein1996growing}
-
 Maybe it's because states are not unitary. Not necessariy requirement: e.g.,
@@
  - bigger picture, world politics as complex system
- - look at Axtell \citep{axtell2000agents}
+ -
 
@@
 There is reason to believe that the orthodox multivariate statistical modeling 
-approach employed in \citet{sechser2010army} is insufficient. Several factors point towards 
-agent-based modeling being a more appropriate choice.
+approach employed in \citet{sechser2010army} is insufficient and leads to
+incorrect conclusions.
 
-One, from a stat
+One, from a statistical perspective, the ability of the model in
+\citet{sechser2010army} to produce reliable insight is questionable.
+Observational data. Selection into war and deterrence.
+
+Two, linear model poorly suited to represent the complex interactions. List
+some. GDP/capita to democracy to war to mechanization. Internal politics to
+external politics to mechanization. Emulation and network effects.
+
+Agent-based modeling more appropriate.
+
+look at Axtell \citep{axtell2000agents}
+
+Other studies have already done it.
+
+Methodology studies
+\citep{pepinsky2005agents,de2014agent,
+gunitsky2013complexity,bousquet2011beyond,
+harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
+
+Insurgency studies
+\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
+
+Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
+
+Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
+
+Just because \citep{epstein1996growing}
 

### 2018-07-14 paper2/paper2.tex
justification

@@
 	inter-state dynamics (e.g., conflict) better explain mechanization choices than do
-	intra-state factors (e.g., culture). This paper demonstrates that
-	the multivariate statistical approach employed in \citet{sechser2010army} is
-	inadequate and that agent-based modeling is needed to capture
-	international security competition as a complex system. Candidate
-	agent-based models are presented.
+	intra-state factors (e.g., culture). This paper argues for the
+	necessity of agent-based modeling of mechanization. ``Agentizing'' the
+	model of \citet{sechser2010army} shows that their findings do not hold.
+	An alternate model that treats mechanization as a complex system is presented 
+	and discussed.
 
@@
 
-Why ABM:
- - inadequacy of SS - VIF, linear model
- - uncertainty of causation
- - endogeneity
- - GDP / capita explains this...but also everything else
- - have to go with simulation
- - bigger picture, world politics as complex system
- -
-
-Things they don't capture: selection into war and deterrence, gdp / capita to
-political structure, institutional culture and choice, network effects and
-emulation
-
-Two central claims of this paper are that the multivariate statistical approach
-of \citet{sechser2010army} is inadequate and agent-based modeling is needed.
-They are addressed in turn.
-
-\citet{sechser2010army} gather observational panel data that consists of a measure 
-of military mechanization $y$ indexed by state $i$ and observation time $k$, as well as
-time-indexed covariates $x$. The association of covariates with mechanization is
-estimated via ordinary least squares (OLS) linear regression:
-
-\begin{equation}
-	y_{ik} = \beta_0 + \beta_j x_{ijk} + \epsilon \hspace{1em}
-\end{equation}
-
-Multicollinearity - everything correlated with each other.
-
-The fitted model attempts to characterize the separate
-effect of each factor upon mechanization. However, evidence suggests that
-the covariates in fact are strongly cross-correlated. The problem of
-multi-collinearity can be expressed both
-
 \section{Introduction}
@@
 reviews the academic literature on military mechanization. It discusses why
-orthodox modeling strategies fail to capture international security competition
+orthodox modeling strategies fail to fully describe international security competition
 as a complex system and why agent-based modeling is needed. In support of these
-claims, it ``agentizes'' the leading model of military mechanization and shows that the
-results of doing cast doubt on it. Finally, it presents a
+claims, it ``agentizes'' the leading model of military mechanization and shows
+that it does not hold. Finally, it presents a
 novel agent-based model that better represents military mechanization.
-
-In summary, the orthodox view of mechanization is that it is a
-reflexive process: states choose their mechanization posture largely based on
-that neighbors, enemies, and recent military experiences (e.g. defeat in
-counter-insurgency). This paper argues
-that this view---based on traditional longitudinal data analysis---is ripe for
-agentization. It implicitly argues for autonomous agents (states) that have
-cognition regarding mechanization that takes into account other agents (e.g.
-enemies), spatiality (e.g. neighbors), and learning from experience. This paper
-builds such a model and validates it against the empirical record, using the
-National Mechanization Index of \citet{sechser2010army}. It finds that the 
-orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
-causes states to
-mechanize much faster and to a greater degree than occurs in the real-world.
-This implies the existence of some countervailing factor (possibly a negative
-feedback loop) that mitigates the extent of mechanization. Candidates for what
-this could be mostly center on ``internal'' factors---e.g., culture, economy,
-and so on---as opposed to the ``external'', security environment-centric
-orthodox view.
 
@@
 
-There are many candidates for the causes of mechanization. Inter-state
-competition and war is one. States may choose choose mechanization policies 
+There are many potential causes of mechanization. Inter-state
+competition and war is a prominent candidate. States may choose choose mechanization policies 
 that they believe will allow them to prevail against potential adversaries, or to send a
@@
 assemble a longitudinal dataset at the country-year level of the mechanization
-of all states between 1979 and 2001. They conduct regression analysis to identify
-the effect of covariates on mechanization. They conclude that out of the
+of all states between 1979 and 2001. They conduct regression analysis on this
+panel data. They use a ordinary least squares (OLS) linear regression model:
+
+\begin{equation}
+y_{ij} = \beta_0 + \sum_{k=1}^P \beta_{ijk}x_{ijk} + \epsilon
+\end{equation}
+
+with $y_{ij}$ the mechanization level of state $i$ in year $j$ (measured as the 
+ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
+x_j \ldots x_{P}$ as covariates hypothesized to be associated with
+mechanization. They conclude that out of the
 aforementioned potential causes---security environment, economy, domestic
@@
 
-\subsection{The case for agent-based modeling}
+\section{The case for agent-based modeling}
 
-There is reason to believe that the orthodox multivariate statistical modeling 
-approach employed in \citet{sechser2010army} is insufficient and leads to
-incorrect conclusions.
+Although \citet{sechser2010army} is a valuable contribution to the literature,
+its orthodox multivariate statistical modeling is insufficient. Agent-based
+modeling is needed to capture mechanization as a complex system. These two
+claims are addressed in turn.
 
-One, from a statistical perspective, the ability of the model in
-\citet{sechser2010army} to produce reliable insight is questionable.
+\subsection{Inadequacy of statistical modeling}
+
+\citet{sechser2010army} operate on observational rather than experimental
+data. The assignment of treatments to subjects is not randomized, causing
+\textit{selection bias}. For example, \citet{sechser2010army} operationalize
+dyadic ``rivalry'' as whether two states recently engaged in a Militarized
+Interstate Dispute (MID) \citep{cow_midb}, and find significant association with mechanization
+(states match their mechanization level to that of their enemy). But interstate
+conflict is not randomly assigned; rather, at least one state must select
+themselves into it \citep{fearon2002selection}. States differ in their
+proclivity for conflict initiation based on many factors, e.g. certain kinds of autocracies may be
+especially war-prone \citep{de1995war}. Thus, correlation attributed to rivalry is in fact
+bundling in many uncontrolled-for covariates, casting doubt on the claimed
+finding. Natural experiments---a situation in which assignation of treatment is
+known to be random---are the typical econometric strategy to avoid bias. For
+example, \citet{angrist2002vouchers} exploit the fact that private school
+vouchers were distributed by lottery in Colombia to obtain unbiased estimates of
+the effect of schooling type (private vs. public) on educational achievement.
+But such situations are rare in international relations---e.g., one cannot
+even conceive of a way in which GDP per capita could be randomly assigned---and so
+\cite{sechser2010army} lack such a control. Given rife selection effects and few
+natural experiments, econometric modeling simply may not be an appropriate tool for mechanization.
+
+The approach of \citet{sechser2010army} does not adequately represent complex interactions. First consider 
+\textit{system decomposibility}. The additive linear model in (1) implicitly assumes 
+that each covariate $x$ is separable and interpretation of the corresponding
+coefficient $\beta$ depends on independence (i.e. a lack of multicollinearity).
+It also assumes that mechanization $y$ is a function of those 
+$x$. But evidence contradicts both assumptions. Take as an example GDP per
+capita, which \citet{sechser2010army} find to be negatively correlated with
+mechanization at the $p < 0.05$ level. We can represent this finding as a causal
+diagram in the style of \citet{pearl2009causality}.\footnote{To be clear,
+\citet{sechser2010army} studiously avoid claiming causation in their text. But
+their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
+direction of association.}
+
+But, a vast body of literature suggests that wealth is associated with
+democracy. Causality appears to flow in both directions, i.e. democracy causes wealth and wealth also
+causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
+Democracy affects conflict propensity \citep{maoz1993normative,russett1994grasping}. 
+Conflict also affects democracy, e.g. a foreign-imposed regime change a la 1956
+Hungary or 2003 Iraq
+\citep{werner1996absolute,lo2008ensuring,downes2013forced}, and mechanization, 
+as claimed by \citet{sechser2010army}. Mechanization affects battlefield success in war
+\citep{biddle2004military,lyall2009rage}. War affects economic growth: 
+consider the contrasting examples of WWII-era United States and present-day
+Syria. If we represent these claims---which, it must be emphasized, do not even
+scratch the surface of the potential system complexity---as a causal diagram, we
+have:
+
+Wealth is 
+Interactions --> causal arrows, GDP per capita, democracy, war, mechanization
+--> decomposibility of system
+
+Path dependence
+
+Cognition
+
+ositive feedback loop
+
+Path dependence
+
+certain kinds of interactions
+
+\subsection{Unreliable data}
+
+Militarized Interstate Disputes Database (MIDB)
+\citep{cow_midb}
+to produce reliable insight is questionable.
 Observational data. Selection into war and deterrence.
+
+MID data sucks
 
@@
 external politics to mechanization. Emulation and network effects.
+
+Lag effects = path dependence
+
+Seeing what others are doing
 
@@
 
+In summary, the orthodox view of mechanization is that it is a
+reflexive process: states choose their mechanization posture largely based on
+that neighbors, enemies, and recent military experiences (e.g. defeat in
+counter-insurgency). This paper argues
+that this view---based on traditional longitudinal data analysis---is ripe for
+agentization. It implicitly argues for autonomous agents (states) that have
+cognition regarding mechanization that takes into account other agents (e.g.
+enemies), spatiality (e.g. neighbors), and learning from experience. This paper
+builds such a model and validates it against the empirical record, using the
+National Mechanization Index of \citet{sechser2010army}. It finds that the 
+orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
+causes states to
+mechanize much faster and to a greater degree than occurs in the real-world.
+This implies the existence of some countervailing factor (possibly a negative
+feedback loop) that mitigates the extent of mechanization. Candidates for what
+this could be mostly center on ``internal'' factors---e.g., culture, economy,
+and so on---as opposed to the ``external'', security environment-centric
+orthodox view.
 \end{document}

### 2018-07-14 paper2/paper2.tex
werk

--- 2018-07-12 paper2/paper2.tex
@@
 \graphicspath{}
+\usepackage{tikz}
 
@@
 
-\title{An agent-based model of military mechanization}
+\title{Military mechanization as a complex system}
 
@@
 	inter-state dynamics (e.g., conflict) better explain mechanization choices than do
-	intra-state factors (e.g., culture). This paper demonstrates that
-	the multivariate statistical approach employed in \citet{sechser2010army} is
-	inadequate and that agent-based modeling is needed to capture
-	international security competition as a complex system. Candidate
-	agent-based models are presented.
+	intra-state factors (e.g., culture).
 
@@
 \newpage
-
-Methodology studies
-\citep{pepinsky2005agents,de2014agent,
-gunitsky2013complexity,bousquet2011beyond,
-harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
-
-Insurgency studies
-\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
-
-Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
-
-Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
-
-Just because \citep{epstein1996growing}
-
-Maybe it's because states are not unitary. Not necessariy requirement: e.g.,
-household is agent in Schelling model.
-
-Cite Cederman to say why I don't do inheritenace - reification of actor
-
-Why ABM:
- - inadequacy of SS - VIF, linear model
- - uncertainty of causation
- - endogeneity
- - GDP / capita explains this...but also everything else
- - have to go with simulation
- - bigger picture, world politics as complex system
- - look at Axtell \citep{axtell2000agents}
-
-Things they don't capture: selection into war and deterrence, gdp / capita to
-political structure, institutional culture and choice, network effects and
-emulation
-
-Two central claims of this paper are that the multivariate statistical approach
-of \citet{sechser2010army} is inadequate and agent-based modeling is needed.
-They are addressed in turn.
-
-\citet{sechser2010army} gather observational panel data that consists of a measure 
-of military mechanization $y$ indexed by state $i$ and observation time $k$, as well as
-time-indexed covariates $x$. The association of covariates with mechanization is
-estimated via ordinary least squares (OLS) linear regression:
-
-\begin{equation}
-	y_{ik} = \beta_0 + \beta_j x_{ijk} + \epsilon \hspace{1em}
-\end{equation}
-
-Multicollinearity - everything correlated with each other.
-
-The fitted model attempts to characterize the separate
-effect of each factor upon mechanization. However, evidence suggests that
-the covariates in fact are strongly cross-correlated. The problem of
-multi-collinearity can be expressed both
 
@@
 assemble a longitudinal dataset at the country-year level of the mechanization
-of all states between 1979 and 2001. They conduct regression analysis to identify
-the effect of covariates on mechanization. They conclude that out of the
-aforementioned potential causes---security environment, economy, domestic
-politics, and ideology / culture---that the first is most important: ``choices about
+of all states between 1979 and 2001. They conduct regression analysis on this
+panel data. They use a ordinary least squares (OLS) linear regression model with corrected standard
+errors,
+
+\begin{equation}
+y_{ij} = \beta_0 + \sum_{k=1}^P \beta_{ijk}x_{ijk} + \epsilon
+\end{equation}
+
+with $y_{ij}$ the mechanization level of state $i$ in year $j$ (measured as the 
+ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
+x_j \ldots x_{P}$ covariates hypothesized to be associated with
+mechanization. They conclude that, out of the
+aforementioned potential causes, the first mentioned is most important: ``choices about
 mechanization are strongly associated with a state's security environment.''
 
-\subsection{The case for agent-based modeling}
+\section{System complexity and agent-based modeling}
 
-There is reason to believe that the orthodox multivariate statistical modeling 
-approach employed in \citet{sechser2010army} is insufficient. Several factors point towards 
-agent-based modeling being a more appropriate choice.
+Although \citet{sechser2010army} is a valuable contribution to the literature,
+its orthodox multivariate statistical modeling is insufficient. Mechanization is a complex system. 
+Agent-based modeling is a more appropriate paradigm for studying it. These three claims are addressed in turn.
 
-One, from a stat
+\subsection{Inadequacy of statistical modeling}
+
+\citet{sechser2010army} operate on observational rather than experimental
+data. The assignment of treatments to subjects is not randomized, causing
+\textit{selection bias}. For example, \citet{sechser2010army} operationalize
+dyadic ``rivalry'' as whether two states recently engaged in a Militarized
+Interstate Dispute (MID) \citep{cow_midb}, and find significant association with mechanization. 
+But interstate conflict is not a randomly assigned treatment. Rather, at least one state must select
+themselves into it \citep{fearon2002selection}. States differ in their
+proclivity for conflict initiation based on many distinct factors. 
+Thus, correlation attributed to rivalry is in fact
+bundling together many uncontrolled-for covariates, casting doubt on the claimed
+finding. Natural experiments---a situation in which assignation of treatment is
+known to be random---are the typical econometric strategy to avoid bias. For
+example, \citet{angrist2002vouchers} exploit the fact that private school
+vouchers were distributed by lottery in Colombia to obtain unbiased estimates of
+the effect of schooling type (private vs. public) on educational achievement.
+But such situations are rare in international relations---e.g., random assignation of war is an
+absurd proposition---and so
+\citet{sechser2010army} lack such a control. Given rife selection effects and few
+natural experiments, econometric modeling alone is an insufficient tool for studying mechanization.
+
+\subsection{System complexity}
+
+\subsubsection{Decomposibility}
+
+The additive linear model in (1) assumes that each covariate $x$ is separable.  
+Interpretation of the corresponding coefficient $\beta$ requires an assumption of independence 
+(i.e. a lack of multicollinearity). It also makes an implicit claim about the direction of
+causation: that mechanization $y$ is a function of those 
+$x$.\footnote{To be clear,
+\citet{sechser2010army} studiously avoid claiming causation in their text. But
+their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
+direction of association.} These assumptions do not hold.
+
+Let us take as an example GDP per capita. \citet{sechser2010army} find it negatively correlated with
+mechanization at the $p < 0.05$ level. Unbiased estimation of this individual coefficient
+requires lack of correlation between GDP per capita and other covariates, either individually or
+jointly. Put differently, \textit{it requires that the system can be neatly decomposed into constituent parts}. Figure 1 is a
+causal diagram depicting this logic. 
+\begin{figure}[h]
+	\centering
+	\caption{Decomposable system}
+	\begin{tikzpicture}
+		\node[draw, rectangle](Mechanization)  at  (0,0) {Mechanization};	
+		\node[draw, rectangle](GDP per capita)  at  (4,0) {GDP per capita};
+		\node[draw, rectangle](Other covariates)  at  (4,-2) {Other covariates};
+		\draw[->] (GDP per capita) -- (Mechanization);
+		\draw[->] (Other covariates) -- (Mechanization);
+	\end{tikzpicture}
+\end{figure}
+
+By standard statistical tests, the assumption of decomposability appears supported. The model's variance
+inflation factor generally does not rise above 2, indicating a lack of
+multicollinearity.\footnote{The high correlation between dummy variables}
+
+However, the literature suggests that (1) is not a correctly specified model. GDP per capita is associated with
+regime type (e.g., democracy vs. autocracy), and causality appears to flow in both directions, i.e. democracy
+causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
+Regime type affects the probability of war: the closest thing to an empirical law in international relations is
+that democracies do not fight each other
+\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. Regime type is widely argued to influence
+battlefield performance in war because it affects the quality of military institutions and hence
+capability of fielded forces \citep{reiter2002democracies,weeks2014dictators,talmadge2015dictator}. 
+Battlefield performance affects regime type, e.g. a losing side may be subject to foreign-imposed
+regime change a la 1939 France or 2003 Iraq
+\citep{werner1996absolute,lo2008ensuring,downes2013forced}. \citet{sechser2010army} claim that the observed
+historical probability of conflict (``rivalry'', in their terminology) affects mechanization; and claim that observed
+battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects the
+probability of war, as it sends some kind of deterrent signal. Mechanization also surely affects
+battlefield performance in war (if it did not, we would not care about it).
+Mechanization also may affect GDP per capita in that states may over-spend on defense and 
+impoverish the country, with the Soviet Union as a prominent example. Deterrence theory requires that 
+battlefield performance affect the probability of war, i.e. militarily 
+potent states are unlikely to be attacked. The exact truth value of these claimed relationships is
+uncertain---in no small part due to the limitations of econometrics that were just addressed---but
+they are generally uncontroversial statements in their respective fields.
+
+\begin{figure}[h]
+	\centering
+	\caption{Complex system}
+	\begin{tikzpicture}
+		\node[draw, rectangle](Mechanization)            at  (0,0)  {Mechanization};	
+		\node[draw, rectangle](GDP per capita)           at  (4,0)  {GDP per capita};
+		\node[draw, rectangle](Democracy)                at  (8,0) {Democracy};
+		\node[draw, rectangle](Probability of war)       at  (0,-4)  {Probability of war};	
+		\node[draw, rectangle](Battlefield performance)  at  (8,-4)  {Battlefield performance};	
+		\draw[->] (Mechanization) to (GDP per capita);
+		\draw[->] (GDP per capita) to (Democracy);
+		\draw[->] (Democracy) to (Battlefield performance);
+		\draw[->] (Battlefield performance) to (Probability of war);
+		\draw[->] (Probability of war) to (Mechanization);
+		\draw[->] (GDP per capita) to[out=90,in=90] (Mechanization);
+		\draw[->] (Democracy) to[out=270,in=270] (GDP per capita);
+		\draw[->] (Battlefield performance) to[out=360,in=360] (Democracy);
+		\draw[->] (Mechanization) to (Battlefield performance);
+		\draw[->] (Mechanization) to[out=180,in=180] (Probability of war);
+		\draw[->] (Democracy) to (Probability of war);
+	\end{tikzpicture}
+\end{figure}
+
+Figure 2 is the logic diagram resulting from the above. It must be emphasized that we are only scratching the 
+surface of system complexity. For example, Figure 2 does not address the nature of relationships 
+between covariates: linear or non-linear? subject to multi-way
+interactions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
+not tractable to decomposition}. Ascertaining cause and effect in this system through simple observation is 
+near-impossible. 
+One instead must perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
+of the nodes and observe the
+resultant system state. But, as previously covered, doing so in international relations is an
+absurd proposition: one cannot mete out international conflict in a real-world experimental
+setting to assess the effect on mechanization. A modeling paradigm that would let us do so is needed.
+
+\subsubsection{Cognition}
+
+There is no law of nature dictating that an increase in one state's mechanization causes an adversary to 
+raise its own mechanization. It also cannot be claimed that many individual choices aggregate into something 
+akin to natural law---the sort of logic used 
+in economics to justify treating supply and demand in a physics-esque manner---because the number of
+people that decide mechanization choices is small. In a personalistic
+dictatorship, it might be $n=1$; even in polities with more dispersed authority, such decisions
+typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
+\textit{A state's mechanization level changes because someone decides to change it}.
+
+If one takes the previous statement to be true, then a model of military mechanization must be in
+some sense a cognitive model. Some agent or set of agents perceives the outside world;
+
+\subsection{The utility of agent-based modeling}
+
+Good for insurgency because its hidden, and
+
+In the typology of Axtell \citep{axtell2000agents}
+
+ABM is useful for exactly such settings.
+
+Methodology studies
+\citep{pepinsky2005agents,de2014agent,
+gunitsky2013complexity,bousquet2011beyond,
+harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
+
+Insurgency studies
+\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
+
+Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
+
+Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
+
+Just because \citep{epstein1996growing}
 

### 2018-07-15 paper2/paper2.tex
onward and upward

@@
 	Military mechanization is the degree to which a state's army is comprised of
-	ground combat vehicles as opposed to infantry. Mechanization is both a 
-	theoretically important topic in international relations scholarship and a practical 
-	concern for policymakers. \citet{sechser2010army} influentially argue that
-	inter-state dynamics (e.g., conflict) better explain mechanization choices than do
-	intra-state factors (e.g., culture).
+	ground combat vehicles as opposed to infantry. Mechanization is an important topic 
+	in international relations scholarship and a practical concern for
+	policymakers. This paper makes novel theoretical contributions by
+	re-envisioning mechanization as a complex system. It also implements an
+	agent-based model of military mechanization. The results of this model
+	challenge the conventional wisdom that mechanization is primarily driven
+	by external factors such as security environment. They imply that
+	some internal factors (e.g. military culture or economic constraints)
+	act as a ``brake'' on mechanization.
 
@@
 
-This paper presents an agent-based model of military mechanization. It
-reviews the academic literature on military mechanization. It discusses why
-orthodox modeling strategies fail to fully describe international security competition
-as a complex system and why agent-based modeling is needed. In support of these
-claims, it ``agentizes'' the leading model of military mechanization and shows
-that it does not hold. Finally, it presents a
-novel agent-based model that better represents military mechanization.
+This paper is divided into five sections. This \textbf{Introduction} gives
+background on military mechanization. The \textbf{Theory} section rigorously
+develops mechanization as a complex system best represented by
+agent-based modeling. The \textbf{Model} section presents a hypothesis and
+implements an agent-based model in Python to test that hypothesis.
+\textbf{Results} are shown. Implications are explored in \textbf{Discussion}.
 
@@
 than labor-rich states \citep{gartzke2001democracy}. Domestic political
-institutions may matter: e.g., autocracies may be more willing than democracies to adopt
+institutions may matter for military composition: e.g., autocracies may be more willing than democracies to adopt
 low-mechanization ``human wave'' tactics
@@
 objectives, which may or may not be coincident with rational strategy.
-Consider Ireland: between the 1920s and 1940s, it seriously considered expending a 
-large portion of its defense budget on building a relatively small number of tanks. 
-They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
-and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
-but tanks were seen as a prestigious signifier of a ``real'' professional
-military and hence desirable
-\citep{farrell1998professionalization,farrell2001transnational}.
 
@@
 assemble a longitudinal dataset at the country-year level of the mechanization
-of all states between 1979 and 2001. They conduct regression analysis on this
+of most states between 1979 and 2001. They conduct regression analysis on this
 panel data. They use a ordinary least squares (OLS) linear regression model with corrected standard
@@
 mechanization. They conclude that, out of the
-aforementioned potential causes, the first mentioned is most important: ``choices about
+aforementioned potential causes, ``choices about
 mechanization are strongly associated with a state's security environment.''
+Their dataset and conclusions have guided further scholarship on military
+mechanization, e.g. \citet{pilster2012democracies}, which mostly follow the same
+statistical approach.
 
-\section{System complexity and agent-based modeling}
+\newpage
 
-Although \citet{sechser2010army} is a valuable contribution to the literature,
-its orthodox multivariate statistical modeling is insufficient. Mechanization is a complex system. 
-Agent-based modeling is a more appropriate paradigm for studying it. These three claims are addressed in turn.
+\section{Theory}
+
+Definition of complex system. 
+The orthodox multivariate statistical modeling approach is insufficient because
+mechanization is a complex system. Agent-based modeling (ABM) is a more 
+appropriate paradigm for studying it. These three
+claims---insufficiency, complexity, and ABM---are addressed in turn.
 
@@
 
-\citet{sechser2010army} operate on observational rather than experimental
+\citet{sechser2010army} use observational rather than experimental
 data. The assignment of treatments to subjects is not randomized, causing
-\textit{selection bias}. For example, \citet{sechser2010army} operationalize
+\textit{selection bias}. For example, they operationalize
 dyadic ``rivalry'' as whether two states recently engaged in a Militarized
-Interstate Dispute (MID) \citep{cow_midb}, and find significant association with mechanization. 
+Interstate Dispute (MID), and find significant association with mechanization. 
 But interstate conflict is not a randomly assigned treatment. Rather, at least one state must select
-themselves into it \citep{fearon2002selection}. States differ in their
-proclivity for conflict initiation based on many distinct factors. 
-Thus, correlation attributed to rivalry is in fact
-bundling together many uncontrolled-for covariates, casting doubt on the claimed
-finding. Natural experiments---a situation in which assignation of treatment is
-known to be random---are the typical econometric strategy to avoid bias. For
+itself into it \citep{fearon2002selection}. States differ in their
+proclivity for conflict, based on many factors. 
+Thus, ``rivalry'' is in fact
+bundling together many uncontrolled covariates, casting doubt on the claimed
+finding. Similar problems arise for other variables included in the model.
+Natural experiments---a situation in which treatment is
+known to be randomly assigned---are the typical econometric strategy to avoid selection bias. For
 example, \citet{angrist2002vouchers} exploit the fact that private school
@@
 the effect of schooling type (private vs. public) on educational achievement.
-But such situations are rare in international relations---e.g., random assignation of war is an
-absurd proposition---and so
-\citet{sechser2010army} lack such a control. Given rife selection effects and few
-natural experiments, econometric modeling alone is an insufficient tool for studying mechanization.
+But such situations are rare in IR: e.g., random assignment of war is an
+absurd proposition. \citet{sechser2010army} lack a good control
+because it probably does not exist. Reliance on selection-riddled
+observational data means that econometrics are at best an incomplete
+tool for studying mechanization.
 
@@
 
-The additive linear model in (1) assumes that each covariate $x$ is separable.  
-Interpretation of the corresponding coefficient $\beta$ requires an assumption of independence 
-(i.e. a lack of multicollinearity). It also makes an implicit claim about the direction of
-causation: that mechanization $y$ is a function of those 
-$x$.\footnote{To be clear,
-\citet{sechser2010army} studiously avoid claiming causation in their text. But
+The additive linear model in (1) assumes that each covariate $x$ is separable. Take as an example 
+GDP per capita.\footnote{The example works as well with any
+other covariate.} \citet{sechser2010army} find it negatively related with mechanization at the $p < 0.05$ level. 
+But a valid interpretation requires an assumption
+of its independence from other covariates (i.e. no multicollinearity).
+It also makes an implicit claim about causal direction: that
+mechanization $y$ is a function of various
+$x$, and not the other way around.\footnote{To be clear,
+\citet{sechser2010army} do not claim causation in their text. But
 their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
-direction of association.} These assumptions do not hold.
-
-Let us take as an example GDP per capita. \citet{sechser2010army} find it negatively correlated with
-mechanization at the $p < 0.05$ level. Unbiased estimation of this individual coefficient
-requires lack of correlation between GDP per capita and other covariates, either individually or
-jointly. Put differently, \textit{it requires that the system can be neatly decomposed into constituent parts}. Figure 1 is a
-causal diagram depicting this logic. 
-\begin{figure}[h]
+direction of association.} Put differently, \textit{it requires that the
+observed system can be neatly 
+decomposed into constituent parts}. Figure 1 is a causal diagram depicting this logic. 
+\begin{figure}[h!]
 	\centering
@@
 
-By standard statistical tests, the assumption of decomposability appears supported. The model's variance
-inflation factor generally does not rise above 2, indicating a lack of
-multicollinearity.\footnote{The high correlation between dummy variables}
+Initial statistical testing seems to support this assumption. The model's variance
+inflation factor (VIF) does not indicate multicollinearity.\footnote{See the Appendix for
+a supporting table. Some coefficients have VIF $> 8$, which strongly
+indicates multicollinearity. However, this only occurs with ``dummy'' variables
+and the variable to which they relate, and so is not problematic.} However, VIF
+is contingent on the specific model being tested, and evidence suggests that 
+(1) is not a correctly specified model. It over-simplifies or omits key
+relationships.
 
-However, the literature suggests that (1) is not a correctly specified model. GDP per capita is associated with
-regime type (e.g., democracy vs. autocracy), and causality appears to flow in both directions, i.e. democracy
-causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
+A large literature suggests that GDP per capita is associated with
+regime type (e.g., democracy vs. autocracy), and causality appears to flow in
+both directions. Democracy causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
 Regime type affects the probability of war: the closest thing to an empirical law in international relations is
 that democracies do not fight each other
-\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. Regime type is widely argued to influence
+\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. 
+Regime type is widely argued to influence
 battlefield performance in war because it affects the quality of military institutions and hence
@@
 historical probability of conflict (``rivalry'', in their terminology) affects mechanization; and claim that observed
-battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects the
-probability of war, as it sends some kind of deterrent signal. Mechanization also surely affects
-battlefield performance in war (if it did not, we would not care about it).
-Mechanization also may affect GDP per capita in that states may over-spend on defense and 
-impoverish the country, with the Soviet Union as a prominent example. Deterrence theory requires that 
-battlefield performance affect the probability of war, i.e. militarily 
+battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects
+battlefield performance in war; if it did not, we would not care about it
+\citep[see also][]{lyall2009rage}.
+Mechanization may affect GDP per capita in that states can over-spend on defense and 
+impoverish the country
+\citep{cappelen1984military,ward1992sizing,abu2003government}. Deterrence theory requires that 
+battlefield performance affect the probability of war in that militarily 
 potent states are unlikely to be attacked. The exact truth value of these claimed relationships is
@@
 between covariates: linear or non-linear? subject to multi-way
-interactions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
-not tractable to decomposition}. Ascertaining cause and effect in this system through simple observation is 
-near-impossible. 
-One instead must perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
-of the nodes and observe the
-resultant system state. But, as previously covered, doing so in international relations is an
-absurd proposition: one cannot mete out international conflict in a real-world experimental
-setting to assess the effect on mechanization. A modeling paradigm that would let us do so is needed.
+interactions? necessity and sufficiency distinctions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
+not tractable to decomposition}. Ascertaining cause and effect in this system
+through simple observation is impossible because of endogeneity. Moreover, controlling for
+confounding variables breaks down when the controlled-for variable directly
+relates to the outcome of interest \citep[``post-treatment bias'',
+c.f.][]{king2010hard}. One would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
+of the nodes and observe the resultant system state. But, as previously covered, one generally cannot conduct
+experimental trials in international relations.  A modeling paradigm that would let us do so is needed.
 
@@
 There is no law of nature dictating that an increase in one state's mechanization causes an adversary to 
-raise its own mechanization. It also cannot be claimed that many individual choices aggregate into something 
-akin to natural law---the sort of logic used 
+raise its own mechanization. It also is not the case that many individual choices aggregate into something 
+akin to natural law---the logic used 
 in economics to justify treating supply and demand in a physics-esque manner---because the number of
 people that decide mechanization choices is small. In a personalistic
-dictatorship, it might be $n=1$; even in polities with more dispersed authority, such decisions
+dictatorship, it might be $n=1$. Even in polities with more dispersed authority, such decisions
 typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
@@
 
-If one takes the previous statement to be true, then a model of military mechanization must be in
-some sense a cognitive model. Some agent or set of agents perceives the outside world;
+If one takes the previous statement to be true, then a model of military
+mechanization must be at least partly a cognitive model. Some agent or set of agents perceives the outside
+world; processes that information; and takes action. Orthodox studies of
+mechanization incorporate cognition to some degree. For example,
+\citet{sechser2010army} test for learning: whether states that fail in
+counter-insurgency wars decrease their mechanization as a result of this
+``anti-mechanization lesson.'' This depiction, however, fails to capture 
+key phenomena.
+
+Human beings learn not only from their own experiences, but also from observing others.
+State decisionmakers are no exception. For example, the stunning success of
+Coalition precision-guided munitions in the 1991 Gulf War led other states who
+were not direct combatants to reform their own militaries
+\citep{scobell2011chinese}. There is no guarantee that such emulation always
+works out. Consider Ireland: between the 1920s and 1940s, it sought to spend a
+large portion of its defense budget on building a small number of
+tanks, which at that time were novel technology being implemented by the leading
+European militaries.
+They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
+and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
+but tanks were seen as a prestigious signifier of a ``real'' professional
+military and hence desirable
+\citep{farrell1998professionalization,farrell2001transnational}. Such learning
+is absent from orthodox models of military mechanization.
 
@@
 
-Good for insurgency because its hidden, and
+In light of the above, agent-based modeling (ABM) is 
+uniquely well-suited to study mechanization. Computer simulations provide the possibility 
+to conduct controlled experiments \textit{in silico} \citep{epstein1996growing} that are not possible 
+in the real world, freeing us from the tyranny of selection bias. Because the
+modeler can adjust various ``dials'' and observe the outcome, ABM allows causal
+inference or at least insight even in highly complex, perhaps unsolvable systems
+\citep[c.f.][]{axtell2000agents}. ABM focuses on autonomous agents with
+cognition, for which even relatively simple behavioral rules can generate
+behavior at least as complex as mechanization
+\citep{gilbert2005simulation,miller2009complex}.
 
-In the typology of Axtell \citep{axtell2000agents}
+Complex systems and ABM have been used in IR. Many methodology-focused
+works have identified their utility to the field \citep[some prominent examples
+are][]{modelski1990world,jervis1998system,pepinsky2005agents,de2014agent}.
+Although actual usage has somewhat lagged behind theoretical enthusiasm,
+ABM has been used to explore emergent 
+balance of power \citep{cederman1997emergent}, international negotiation
+\citep{earnest2008coordination}, and international crises and conflict
+\citep{masad2016dissertation}. It is especially popular for studying
+population-centric security phenomena such as terrorism and insurgency
+\citep{cioffi2010mason,bhavnani2012modeling,weidmann2013violence,bhavnani2014group}.
 
-ABM is useful for exactly such settings.
-
-Methodology studies
-\citep{pepinsky2005agents,de2014agent,
-gunitsky2013complexity,bousquet2011beyond,
-harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}
-
-Insurgency studies
-\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}
-
-Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}
-
-Fundamentals, e.g. cooperation \citep{axelrod1981evolution}
-
-Just because \citep{epstein1996growing}
-
-are ripe for agentization. The
-conclusions drawn implicitly claim
-that states are autonomous agents; that they have
-cognition regarding how to perceive the outside world
-and how to react to it; that spatiality and
-neighborhood matter; that inter-agent relationships matter; and that
-agents learn from experience. Agent-based modeling is uniquely well-suited to
-represent such phenomena \citep{gilbert2005simulation,miller2009complex}.
-
-Importantly, the findings of \citet{sechser2010army} imply a positive feedback
-loop. State A has high
-mechanization, causing enemy State B to raise their mechanization, causing State
-A to raise their mechanization... and so on. Such feedback loops are a hallmark
-of complex systems and again are well-suited for representation by agent-based
-modeling. This specific feedback loop is often called
-the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
-of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.
+The author's sense is that adoption of ABM in IR has lagged because it is
+regarded as non-empirical. ABM offers no guarantees: just
+because one's model ``works'' (internally consistent, externally plausible) 
+does not mean it is true. Econometrics are the dominant methodological approach 
+in quantitative IR because they seem to offer an objective yardstick for competing hypotheses: 
+does A or B best fit the data? This paper contends that
+epistemological limitations (most importantly, selection bias stemming from
+non-randomization, and system complexity) mean that econometrics do not
+provide reliable statistical inference for certain classes of IR problems, among
+them mechanization \citep[c.f.][for a concurring opinion]{king2010hard}. 
+If neither econometrics nor ABM can offer 
+certitude that a hypothesis is true---as is the case for mechanization---then the latter 
+should be preferred for its advantages in experimentation, transparency, researcher control, 
+and amenability to complex systems.
 
@@
 phenomena despite some early work on this front
-\citep[see][]{cederman1997emergent,axelrod1997complexity}. 
+\citep[see][]{,axelrod1997complexity}. 
 This paper's agentization of an orthodox security
@@
 orthodox view.
+
+are ripe for agentization. The
+conclusions drawn implicitly claim
+that states are autonomous agents; that they have
+cognition regarding how to perceive the outside world
+and how to react to it; that spatiality and
+neighborhood matter; that inter-agent relationships matter; and that
+agents learn from experience. Agent-based modeling is uniquely well-suited to
+represent such phenomena .
+
+Importantly, the findings of \citet{sechser2010army} imply a positive feedback
+loop. State A has high
+mechanization, causing enemy State B to raise their mechanization, causing State
+A to raise their mechanization... and so on. Such feedback loops are a hallmark
+of complex systems and again are well-suited for representation by agent-based
+modeling. This specific feedback loop is often called
+the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
+of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{l r}
+		\hline
+		\textbf{Variable} & \textbf{VIF} \\ 
+		\hline
+		log\_mech\_enemies & 8.76 \\ 
+		enemies\_dummy & 9.12 \\ 
+		log\_mech\_neighbors & 2.19 \\ 
+		neighbors\_dummy & 2.03 \\ 
+		log\_mountainous & 1.26 \\ 
+		antimech\_lesson\_2 & 1.21 \\ 
+		promech\_lesson\_2 & 1.11 \\ 
+		democracy & 1.69 \\ 
+		military\_govt & 1.13 \\ 
+		instability & 1.12 \\ 
+		log\_gdp\_cap & 2.10 \\ 
+		iron\_steel & 1.32 \\ 
+		log\_mech\_allies & 12.64 \\ 
+		allies\_dummy & 12.59 \\ 
+		cold\_war\_bloc & 1.48 \\ 
+		igos & 2.03 \\ 
+		post\_soviet & 1.33 \\ 
+		log\_mech\_lag2 & 1.57 \\ 
+		\hline
+	\end{tabular}
+\end{table}
 \end{document}

### 2018-07-15 paper2/paper2.tex
good

@@
 
-\title{Military mechanization as a complex system}
+\title{Military mechanization as a complex adaptive system}
 
@@
 	policymakers. This paper makes novel theoretical contributions by
-	re-envisioning mechanization as a complex system. It also implements an
+	re-envisioning mechanization as a complex adaptive system. It implements an
 	agent-based model of military mechanization. The results of this model
 	challenge the conventional wisdom that mechanization is primarily driven
-	by external factors such as security environment. They imply that
-	some internal factors (e.g. military culture or economic constraints)
-	act as a ``brake'' on mechanization.
+	by external factors such as security environment.
 
@@
 background on military mechanization. The \textbf{Theory} section rigorously
-develops mechanization as a complex system best represented by
+develops mechanization as a complex adaptive system best represented by
 agent-based modeling. The \textbf{Model} section presents a hypothesis and
-implements an agent-based model in Python to test that hypothesis.
+implements an agent-based model to test that hypothesis.
 \textbf{Results} are shown. Implications are explored in \textbf{Discussion}.
@@
 
-States differ in the composition of their ground forces.\footnote{States may have multiple organizations that
+States differ in the composition of their ground forces.\footnote{States may 
+have multiple organizations that
 conduct ground combat operations: e.g., the United States operates both an Army
@@
 mechanization. They conclude that, out of the
-aforementioned potential causes, ``choices about
+potential causes mentioned above, ``choices about
 mechanization are strongly associated with a state's security environment.''
-Their dataset and conclusions have guided further scholarship on military
+States that have engaged in conflict with a highly-mechanized adversary and states
+with highly mechanized neighbors tend to higher mechanization; conversely, states facing
+domestic insurgency or that have lost counter-insurgency wars tend to have lower
+mechanization rates. Their dataset and conclusions have guided further scholarship on military
 mechanization, e.g. \citet{pilster2012democracies}, which mostly follow the same
@@
 
-\newpage
-
 \section{Theory}
 
-Definition of complex system. 
 The orthodox multivariate statistical modeling approach is insufficient because
-mechanization is a complex system. Agent-based modeling (ABM) is a more 
+mechanization is a complex adaptive system. Agent-based modeling (ABM) is a more 
 appropriate paradigm for studying it. These three
@@
 Natural experiments---a situation in which treatment is
-known to be randomly assigned---are the typical econometric strategy to avoid selection bias. For
+known to be randomly assigned---are the ideal strategy to avoid selection bias. For
 example, \citet{angrist2002vouchers} exploit the fact that private school
@@
 But such situations are rare in IR: e.g., random assignment of war is an
-absurd proposition. \citet{sechser2010army} lack a good control
-because it probably does not exist. Reliance on selection-riddled
-observational data means that econometrics are at best an incomplete
-tool for studying mechanization.
+absurd proposition. There are statistical methods for correction of selection
+bias, such as Heckman correction and propensity score matching, but system
+complexity prevents their application, as covered in the following sub-section.
 
-\subsection{System complexity}
+\subsection{Complex adaptive system}
 
-\subsubsection{Decomposibility}
+There is no single universally accepted definition of a complex adaptive system
+\citep{miller2009complex}. However, two core features are
+that the system is non-decomposable, and that the system features cognitive
+agents. The following sections detail how mechanization meets this description.
+
+\subsubsection{Decomposability}
 
@@
 uncertain---in no small part due to the limitations of econometrics that were just addressed---but
-they are generally uncontroversial statements in their respective fields.
+they are generally mainstream statements in their respective fields.
 
@@
 between covariates: linear or non-linear? subject to multi-way
-interactions? necessity and sufficiency distinctions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
-not tractable to decomposition}. Ascertaining cause and effect in this system
-through simple observation is impossible because of endogeneity. Moreover, controlling for
-confounding variables breaks down when the controlled-for variable directly
-relates to the outcome of interest \citep[``post-treatment bias'',
-c.f.][]{king2010hard}. One would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
-of the nodes and observe the resultant system state. But, as previously covered, one generally cannot conduct
+interactions? necessity and sufficiency distinctions? Moreover, it only includes 
+abstract variables and not individual agents (a point addressed in the following
+sub-section). But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
+not tractable to decomposition}. We cannot isolate one component from the
+greater whole.
+
+This system complexity disallows previously-mentioned statistical techniques
+for correction of selection bias. Both Heckman correction and propensity score
+matching boil down to using covariates to predict (and hence control for) the chances 
+of receiving treatment. For example, we might assess the effect of war on
+mechanization by controlling for the factors that affect propensity for war. 
+But these techniques break down if the confounding variables also directly
+relate to the outcome of interest, or if the direction of causality flows both
+wars. Both are true for mechanization 
+(c.f. Figure 2). This ``post-treatment bias'' \citep{king2010hard} means we need
+some technique other than econometrics to perform causal inference regarding
+mechanization.
+
+We would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
+of the nodes and observe the resultant system state. This sort of
+experimentation might allow us observe the system's sensitivity (or lack thereof) to changes
+in input. But, as previously covered, one generally cannot conduct
 experimental trials in international relations.  A modeling paradigm that would let us do so is needed.
 
-\subsubsection{Cognition}
+\subsubsection{Cognitive agents}
 
@@
 typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
-\textit{A state's mechanization level changes because someone decides to change it}.
+\textit{A state's mechanization level changes because someone decides to change it}. If one takes 
+the previous statement to be true, then a model of military mechanization must be a model of 
+cognitive agents.  Some agent or set of agents perceives the outside
+world; processes that information; and takes action.
 
-If one takes the previous statement to be true, then a model of military
-mechanization must be at least partly a cognitive model. Some agent or set of agents perceives the outside
-world; processes that information; and takes action. Orthodox studies of
-mechanization incorporate cognition to some degree. For example,
+Orthodox studies of mechanization incorporate cognition to some degree. For example,
 \citet{sechser2010army} test for learning: whether states that fail in
@@
 tanks, which at that time were novel technology being implemented by the leading
-European militaries.
-They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
+European militaries. They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
 and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
@@
 military and hence desirable
-\citep{farrell1998professionalization,farrell2001transnational}. Such learning
-is absent from orthodox models of military mechanization.
+\citep{farrell1998professionalization,farrell2001transnational}. A more
+sophisticated representation of cognition that incorporates such
+learning-from-others is needed.
 
@@
 behavior at least as complex as mechanization
-\citep{gilbert2005simulation,miller2009complex}.
+\citep{gilbert2005simulation}.
 
@@
 
+\section{Model}
+
+This section of the paper describes the model. It does not use the ODD protocol
+of \citet{grimm2006standard} but covers many of the same points.
+
 \subsection{Research question and strategy}
@@
 and paper, though it is discussed.
-
-\section{Model}
-
-This section of the paper describes the model. It does not use the ODD protocol
-of \citet{grimm2006standard} but covers many of the same points.
-
 \subsection{Agents}

### 2018-07-17 paper2/paper2.tex
recontextualized

@@
 	agent-based model of military mechanization. The results of this model
+	both confirm that mechanization is a complex adaptive system and
 	challenge the conventional wisdom that mechanization is primarily driven
@@
 There is no single universally accepted definition of a complex adaptive system
-\citep{miller2009complex}. However, two core features are
-that the system is non-decomposable, and that the system features cognitive
-agents. The following sections detail how mechanization meets this description.
+\citep{miller2009complex}. However, some core features are
+that the system is non-decomposable; that the system features autonomous, cognitive
+agents; and that the individual behaviors of these agents aggregate upwards into
+complex system behavior (e.g., stability, chaos, feedback loops, and the like). 
+The following sections detail how mechanization meets the first two parts of
+this description (later sections deal with system behavior).
 
@@
 military and hence desirable
-\citep{farrell1998professionalization,farrell2001transnational}. A more
-sophisticated representation of cognition that incorporates such
-learning-from-others is needed.
+\citep{farrell1998professionalization,farrell2001transnational}.
 
@@
 
-This section of the paper describes the model. It does not use the ODD protocol
-of \citet{grimm2006standard} but covers many of the same points.
+This section of the paper describes research hypotheses and an agent-based
+model used to test them.
 
-\subsection{Research question and strategy}
+\subsection{Research hypotheses}
 
@@
 and paper, though it is discussed.
-\subsection{Agents}
+
+\subsection{Model}
+
+\subsubsection{Agents}
 
@@
 
-\subsection{Environment}
+\subsubsection{Environment}
 
@@
 
-\subsection{Cognition}
+\subsubsection{Cognition}
 
@@
 
-\subsection{Time}
+\subsubsection{Time}
 
@@
 
-\subsection{Output Data, Verification, and Validation}
+\subsubsection{Output Data, Verification, and Validation}
 

### 2018-07-18 paper2/paper2.tex
good job tom

@@
 develops mechanization as a complex adaptive system best represented by
-agent-based modeling. The \textbf{Model} section presents a hypothesis and
-implements an agent-based model to test that hypothesis.
+agent-based modeling. The \textbf{Model} section presents hypotheses and
+implements an agent-based model to test them.
 \textbf{Results} are shown. Implications are explored in \textbf{Discussion}.
@@
 ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
-x_j \ldots x_{P}$ covariates hypothesized to be associated with
+x_k \ldots x_{P}$ covariates hypothesized to be associated with
 mechanization. They conclude that, out of the
@@
 mechanization rates. Their dataset and conclusions have guided further scholarship on military
-mechanization, e.g. \citet{pilster2012democracies}, which mostly follow the same
-statistical approach.
+mechanization \citep[e.g.][]{pilster2012democracies},
 
@@
 \citep{miller2009complex}. However, some core features are
-that the system is non-decomposable; that the system features autonomous, cognitive
-agents; and that the individual behaviors of these agents aggregate upwards into
+that the system comprises many constituent parts that interact to form a
+non-decomposable whole; that the system features autonomous, cognitive
+agents; and that the relatively simple interactions of these agents aggregate upwards into
 complex system behavior (e.g., stability, chaos, feedback loops, and the like). 
@@
 surface of system complexity. For example, Figure 2 does not address the nature of relationships 
-between covariates: linear or non-linear? subject to multi-way
-interactions? necessity and sufficiency distinctions? Moreover, it only includes 
+between covariates: linear or non-linear? Subject to multi-way
+interactions? Necessity vs. sufficiency distinctions? Moreover, it only includes 
 abstract variables and not individual agents (a point addressed in the following
@@
 not tractable to decomposition}. We cannot isolate one component from the
-greater whole.
+greater whole. Separating cause from effect in such a system is challenging even
+at the most basic conceptual level
+\citep{pearl2009causality}. It also becomes statistically intractable.
 
-This system complexity disallows previously-mentioned statistical techniques
-for correction of selection bias. Both Heckman correction and propensity score
+This system complexity disallows previously-mentioned
+techniques for mitigating selection bias. Both Heckman correction and propensity score
 matching boil down to using covariates to predict (and hence control for) the chances 
@@
 relate to the outcome of interest, or if the direction of causality flows both
-wars. Both are true for mechanization 
+ways. Both are true for mechanization 
 (c.f. Figure 2). This ``post-treatment bias'' \citep{king2010hard} means we need
@@
 
-We would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
-of the nodes and observe the resultant system state. This sort of
+Complex systems theory \citep[ultimately stemming to][]{simon1996artificial}
+offers a way forward. Although we cannot separate the many parts from the whole,
+we would like to modify the value at one of the nodes and observe the resultant system state. This sort of
 experimentation might allow us observe the system's sensitivity (or lack thereof) to changes
@@
 
-Orthodox studies of mechanization incorporate cognition to some degree. For example,
-\citet{sechser2010army} test for learning: whether states that fail in
-counter-insurgency wars decrease their mechanization as a result of this
-``anti-mechanization lesson.'' This depiction, however, fails to capture 
-key phenomena.
-
-Human beings learn not only from their own experiences, but also from observing others.
+Moreover, human beings learn not only from their own experiences, but also from observing others.
 State decisionmakers are no exception. For example, the stunning success of
@@
 were not direct combatants to reform their own militaries
-\citep{scobell2011chinese}. There is no guarantee that such emulation always
-works out. Consider Ireland: between the 1920s and 1940s, it sought to spend a
+\citep{scobell2011chinese}. There also is no guarantee that such cognition
+follows typical assumptions of bounded rationality. 
+Consider Ireland: between the 1920s and 1940s, it sought to spend a
 large portion of its defense budget on building a small number of
-tanks, which at that time were novel technology being implemented by the leading
-European militaries. They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
-and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
-but tanks were seen as a prestigious signifier of a ``real'' professional
-military and hence desirable
-\citep{farrell1998professionalization,farrell2001transnational}.
+tanks. This was a near-suicidal defense posture, given Ireland's
+conventional weakness relative to the United Kingdom (the main object of Irish
+defense thought). Rationally, Ireland should have pursued a low-mechanization,
+guerrilla defense strategy given their weakness and given that exactly such a
+strategy defeated the British in the 1919-22 war of independence. However,  
+tanks at the time were novel technology being implemented by the leading
+European militaries, and so were seen as a prestigious signifier of a ``real'' professional
+military \citep{farrell1998professionalization,farrell2001transnational}. These examples
+support the larger point that perception and cognition are fundamental to
+understanding mechanization. A modeling framework that explicitly represents these factors
+thus would be useful.
 
@@
 are][]{modelski1990world,jervis1998system,pepinsky2005agents,de2014agent}.
-Although actual usage has somewhat lagged behind theoretical enthusiasm,
-ABM has been used to explore emergent 
+Although actual usage has somewhat lagged behind theoretical
+enthusiasm---a point addressed in more detail below---ABM has been used to explore emergent 
 balance of power \citep{cederman1997emergent}, international negotiation
@@
 regarded as non-empirical. ABM offers no guarantees: just
-because one's model ``works'' (internally consistent, externally plausible) 
-does not mean it is true. Econometrics are the dominant methodological approach 
-in quantitative IR because they seem to offer an objective yardstick for competing hypotheses: 
-does A or B best fit the data? This paper contends that
-epistemological limitations (most importantly, selection bias stemming from
-non-randomization, and system complexity) mean that econometrics do not
-provide reliable statistical inference for certain classes of IR problems, among
-them mechanization \citep[c.f.][for a concurring opinion]{king2010hard}. 
-If neither econometrics nor ABM can offer 
-certitude that a hypothesis is true---as is the case for mechanization---then the latter 
-should be preferred for its advantages in experimentation, transparency, researcher control, 
-and amenability to complex systems.
+because one's model ``works''  does not mean it is true. Econometrics dominate quantitative 
+IR because they offer an objective yardstick for competing hypotheses. 
+This paper has demonstrated that econometrics, although useful in many situation, do not
+provide reliable statistical inference for certain classes of IR problems \citep[c.f.][for a concurring opinion]{king2010hard}. 
+For such cases (and mechanization is one of them), ABM no longer is at the
+disadvantage in causal inference, and has important advantages in experimentation, 
+researcher control, and amenability to complex systems
 
@@
 
-This section of the paper describes research hypotheses and an agent-based
-model used to test them.
+This section of the paper describes the research strategy and specific
+hypotheses, and an agent-based model used to test them.
 
-\subsection{Research hypotheses}
+\subsection{Research strategy and hypotheses}
 
-I hypothesize that the arguments of \citet{sechser2010army} are incomplete and
-that some omitted variable or set of variables acts as a ``brake'' on 
-mechanization.\footnote{By ``omitted'', I mean that Sechser and Saunders conclude that the variable
+This paper's strategy is not to propose a master
+model of mechanization that incorporates all factors raised in Figure 2 and
+the subsequent discussion. Rather, it is to start with a relatively simple
+agent-based model of military mechanization that ``agentizes'' extant
+traditional scholarship on mechanization. This research strategy is deliberate. 
+The goal is to maintain maximum correspondence with extant scholarship on mechanization 
+so that the results of an agentized version of it can be fairly compared.
+
+The primary hypothesis is that the arguments of \citet{sechser2010army}, which
+center on security environment, are incomplete; some omitted variable or set of
+variables must act 
+as a ``brake'' on  mechanization.\footnote{By ``omitted'', I mean that Sechser and Saunders conclude that the variable
 is not very important for mechanization, not that they do not address the
-variable at all.} An agent-based model is used to test this hypothesis. In broad
+variable at all.} If mechanization is externally-driven, then one agent raising their
+mechanization causes another agent to do so, causes another agent to do so\ldots
+and so on. This \textit{positive feedback loop} would reverbate through the
+system and drive mechanization upward. Importantly, it acts constantly, whereas
+the ``anti-mechanization lessons'' identified by \citet{sechser2010army} (e.g.,
+defeat in counter-insurgency) are episodic. This paper thus hypothesizes that a
+faithful agentization of this orthodox model will result in too-high
+mechanization levels as compared to the historical record.
+
+\subsection{Model overview}
+
+An agent-based model is used to test this hypothesis. In broad
 terms, agents (states) are defined to have the cognitive and learning behavior
@@
 real-world data from 1979 and allowed to run until 2001. The micro- and
-macro-outcomes are then validated against real-world 2001 data. If
-\citet{sechser2010army} are correct, the model's 2001 outcomes should match
+macro-outcomes then are validated against real-world 2001 data. If
+orthodox theory is correct, the model's 2001 outcomes should match
 real-world 2001 outcomes. If the model mechanization levels in 2001 are much
 higher, then their claimed behavioral rules are incomplete and fail to include
-some ``brake'' on mechanization, matching my hypothesis. Proving exactly what
-that missing variable or set of variables is is outside the scope of this model
-and paper, though it is discussed.
-
-\subsection{Model}
+some ``brake'' on mechanization, matching the hypothesis.
 
@@
 those with a population less than 750,000). Some other states were excluded; for
-why, see the \textbf{Time}
-section. Figure 1 depicts the set of agents.
+why, see the \textbf{Time} section. Figure 1 depicts the set of agents.
 
@@
 districts \citep{roy1993military}. Such inheritance has
-a strong effect on the mechanization of of successor states. Designing
-complex and state-specific ``rules of inheritance'' was too much for this
-initial effort but is an area for further development.
+a strong effect on the mechanization of of successor states. Building a separate
+sub-model for  complex ``rules of inheritance'' is beyond this initial effort.
 
 \subsubsection{Output Data, Verification, and Validation}
+
+MORE ON VERIFICATION
 
@@
 For some agents, however, mechanization levels are extremely high. 
-Mechanization score is a percentage, and so values over 1 do not make sense, 
+Mechanization score is a ratio, and so values over 1 do not make sense, 
 yet some countries end the simulation with mechanization scores over 1. Note
@@
 
+A parameter sweep reveals that this pattern is relatively insensitive
+to input values. That is to say, changing \textit{NeighborsMechSensitivity}, 
+\textit{EnemiesMechSensitivity}, and \textit{LessonSensitivity} alter the
+magnitude of mechanization but not the overall trend of it being most pronounced
+in states with many neighbors, enemies, and neighbor-enemies. For example, Table
+6 shows the effect of lowering \textit{NeighborsMechSensitivity} and 
+\textit{EnemiesMechSensitivity} to values lower than that used in the baseline
+run. Doing so does not change what state ends up as the most mechanized at
+simulation end, suggesting that the fundamental system behavior is unaltered.
+
+\begin{table}[h]
+	\centering
+	\caption{Parameter sweep}
+	\begin{tabular}{|l r|}
+		\hline
+		\textbf{Parameter Values} & \textbf{Most Mechanized State} \\ 
+		1.0 (baseline) & Israel \\
+		0.75 & Israel \\
+		0.5 & Israel \\
+		0.25 & Israel \\
+		\hline
+	\end{tabular}
+\end{table}
 \section{Discussion}
@@
 diverge from theoretical expectations of the orthodox view.
-
-\subsection{Areas for Improvement}
-
-Having taken a critical view of \citet{sechser2010army}, it only is fair that
-this paper applies that lens to itself. There are two main areas for
-improvement.
-
-One, this model's representation of \citet{sechser2010army} could be tightened
-up, as there currently is divergence between the two. For example, this model
-examines a smaller agent set than their paper. Building ``inheritance''
-processes to allow new states to emerge is desirable. This paper's
-use of statistical coefficients from \citet{sechser2010army} also has some
-methodological problems, a full description of which is omitted here due to
-space constraints. It suffices to say that a better approach would be to conduct
-more extensive parameter sweeps of relevant cognitive factors.
-
-Two, this model could incorporate potential explanations of what the missing
-``brake'' on mechanization might be. Economy and culture could be included in
-the model. Doing so would allow exploration of whether these factors are
-logically consistent and can act as a countervailing factor that controls the
-positive feedback loop of mechanization. It cannot \textit{prove} that these or
-any other factors actually exist in reality, but they can demonstrate how they
-are or are not consistent in a simulated reality.
-
-My hope is that making these changes will make this a publication-quality paper
-capable of serving as one of two required papers for the comprehensive exams.
 

### 2018-07-19 paper2/paper2.tex
more wek

@@
 
-Complex systems theory \citep[ultimately stemming to][]{simon1996artificial}
+Complex systems theory \citep[e.g.][]{simon1996sciences}
 offers a way forward. Although we cannot separate the many parts from the whole,
@@
 those with a population less than 750,000). Some other states were excluded; for
-why, see the \textbf{Time} section. Figure 1 depicts the set of agents.
+why, see the \textbf{Time} section. Figure 3 depicts the set of agents.
 
@@
 predict. Inclusion of other factors not identified as important for
-mechanization---e.g., alliances, economic factors,
-culture, etc.---would be interesting but counter-productive to the research
+mechanization---e.g., alliances, culture, etc.---would be interesting 
+but counter-productive to the research
 question. They are discussed along with this model's findings.
@@
 
+BEEF UP THIS SECTION
+
 Two conclusions should be drawn. One, the model results in much higher
@@
 
+\subsection{Future Research}
+
 \newpage
@@
 \printbibliography[heading=bibnumbered]
-
-In summary, the orthodox view of mechanization is that it is a
-reflexive process: states choose their mechanization posture largely based on
-that neighbors, enemies, and recent military experiences (e.g. defeat in
-counter-insurgency). This paper argues
-that this view---based on traditional longitudinal data analysis---is ripe for
-agentization. It implicitly argues for autonomous agents (states) that have
-cognition regarding mechanization that takes into account other agents (e.g.
-enemies), spatiality (e.g. neighbors), and learning from experience. This paper
-builds such a model and validates it against the empirical record, using the
-National Mechanization Index of \citet{sechser2010army}. It finds that the 
-orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
-causes states to
-mechanize much faster and to a greater degree than occurs in the real-world.
-This implies the existence of some countervailing factor (possibly a negative
-feedback loop) that mitigates the extent of mechanization. Candidates for what
-this could be mostly center on ``internal'' factors---e.g., culture, economy,
-and so on---as opposed to the ``external'', security environment-centric
-orthodox view.
-
-are ripe for agentization. The
-conclusions drawn implicitly claim
-that states are autonomous agents; that they have
-cognition regarding how to perceive the outside world
-and how to react to it; that spatiality and
-neighborhood matter; that inter-agent relationships matter; and that
-agents learn from experience. Agent-based modeling is uniquely well-suited to
-represent such phenomena .
-
-Importantly, the findings of \citet{sechser2010army} imply a positive feedback
-loop. State A has high
-mechanization, causing enemy State B to raise their mechanization, causing State
-A to raise their mechanization... and so on. Such feedback loops are a hallmark
-of complex systems and again are well-suited for representation by agent-based
-modeling. This specific feedback loop is often called
-the ``security dilemma'' and was a main research interest of Thomas Schelling, one of the pioneers
-of agent-based modeling \citep{schelling1960strategy,schelling2006micromotives}.
-
+\newpage
+\appendix
+\centering
+\section*{Appendix A}
 \begin{table}[h]
 	\centering
+	\caption{Variance inflation factor, \citet{sechser2010army}}
 	\begin{tabular}{l r}

### 2018-07-20 paper2/paper2.tex
so close

@@
 
-MORE ON VERIFICATION
-
 The main model output of interest is data on military mechanization. It is provided for
 every agent and every time-step. These agent-level data also can be aggregated
-into system-level measures of military mechanization. Per the research design,
-the intent is to compare the model output to what actually occurred. If the
-model produces military mechanization levels much higher than reality, it means
-that Sechser and Saunder's findings do not stand up to agentization and there is
-some missing ``brake'' on military mechanization. Note that this research
-strategy has validation (comparison of model output to empirical data) at its
-core and so no separate validation is needed. Verification that the model is
-doing what it is intended was accomplished through manual inspection.
+into system-level measures of military mechanization.
+
+Verification--—i.e., making sure that the model does what it is intended to
+do---is a relative activity, not absolute. It is intended to increase confidence in the correctness
+of the model, not to prove that it is perfect \citep{robinson1997simulation}.
+During model development, the model was configured for verbose run-time output. Manual
+inspection of this output confirmed that the model was working as intended. For
+example, output of each actor's \textit{Enemies} at each timestep allowed
+confirmation that the actor was updating its perception in response to scripted
+international conflict in line with the historical record.
+
+Validation---i.e., assessing the correspondence between the model and empirical
+data---is central to the research strategy. The National Mechanization Index
+(NMI) compiled by \citet{sechser2010army} provides empirical data on
+mechanization at the country-year level between 1979 and 2001. In the following
+\textbf{Results} section, model output is compared to this data in
+order to test the hypothesis that the orthodox view of mechanization does
+not stand up to agentization. Because the
+model and empirical data are at the country-year level, validation can take
+place at both the micro-level (i.e., individual state-agents) and at the
+macro-level (i.e., the aggregate mechanization levels of the whole international
+system); and, validation is longitudinal rather than static. 
+Thus, this paper pursues Level 3 validation in the taxonomy of
+\citet{axtell1994understanding}: ``cross-sectional and longitudinal analysis of
+the agent population.''
 
@@
 For some agents, however, mechanization levels are extremely high. 
-Mechanization score is a ratio, and so values over 1 do not make sense, 
+Mechanization score is a ratio, and so values over 1 do not make sense in
+context, 
 yet some countries end the simulation with mechanization scores over 1. Note
@@
 
-BEEF UP THIS SECTION
+BEEF THIS UP AND YOU'RE DONE
 
-Two conclusions should be drawn. One, the model results in much higher
-system-level mechanization than occurred in reality. Two, this higher
-mechanization is not evenly distributed across agents; rather, a small number of
+The model results in much higher
+system-level mechanization than occurred in reality. This higher
+mechanization is not evenly distributed across agents. Rather, a small number of
 agents with many enemies and that all border each other become outliers and
@@
 
-Three, because this model is a
-faithful representation of the factors that \citet{sechser2010army} claim are
+Because this model is a faithful representation of the factors that \citet{sechser2010army} claim are
 most important for mechanization, that model output fails to match reality

### 2018-07-29 paper2/paper2.tex
foo

@@
 
-BEEF THIS UP AND YOU'RE DONE
-
-The model results in much higher
-system-level mechanization than occurred in reality. This higher
+The model results in much higher system-level mechanization than occurred in reality. This higher
 mechanization is not evenly distributed across agents. Rather, a small number of
@@
 mechanization scores in response to each other; hence, a geographic cluster of
-enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their
-mechanization.
+enemies (e.g., the Middle East or Cold War-era Europe) will rapidly increase their mechanization
+and drive system-level mechanization upwards. Modeled sources of anti-mechanization (e.g.,
+defeat in counter-insurgency) are insufficient to hold mechanization to historical levels.
 
-Because this model is a faithful representation of the factors that \citet{sechser2010army} claim are
-most important for mechanization, that model output fails to match reality
-indicates that their claimed model is incomplete. There must be some ``brake''
-on mechanization; put differently, some countervailing factor(s) must mitigate the
-effect of the positive feedback loop identified above. This model cannot
-conclusively identify what that ``brake'' is. The most plausible candidates are
-factors examined by \citet{sechser2010army} and deemed statistically
-insignificant such as regime type and economy, and factors not examined by
-\citet{sechser2010army} at all such as culture.
+Because this model is a faithful representation of the factors that the orthodox
+literature claim are most important for mechanization, that model output fails to match reality
+indicates that the orthodox view of mechanization is incomplete. It must not be
+near-solely driven by external security factors. There must be some additional ``brake''
+on mechanization. Put differently, some countervailing factor(s) must mitigate the
+effect of the positive feedback loop identified above. This model cannot conclusively 
+identify what that ``brake'' is, but preliminary extensions to the model provide
+some insight, as described in the \textbf{Future Research} section.
 
-This paper is a novel contribution to the literature. Although agent-based
-models have found traction in the study of grievance, insurgency, and
-other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
-they quite rarely are used to model international relations and state-centric security
-phenomena despite some early work on this front
-\citep[see][]{,axelrod1997complexity}. 
-This paper's agentization of an orthodox security
-studies model thus is novel in method. Its findings also are important. They
-show that they show a solely ``external'', security
-environment-focused view of mechanization does not stand up to agentization.
-Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
-loops---all hallmarks of a complex system---drive results that significantly
-diverge from theoretical expectations of the orthodox view.
+The model's behavior recreates IR phenomena from the bottom
+up. The aforementioned positive feedback loop is well-studied in IR as the
+``security dilemma'' \citep{jervis1978cooperation}. Note that there is no model
+logic explicitly dictating this behavior. Rather, the model endows agents with 
+only simple rules---e.g., \textit{if my neighbor increases mechanization, 
+then increase my own}---that nevertheless cause the emergence of this behavior.
+That it does so increases confidence in the correctness of the model since
+it in this specific area corresponds to theoretically-expected behavior despite
+not being programmed to do so.
+
+More broadly, model behavior supports the argument that mechanization is
+a complex system. There is no definitive checklist for what constitutes
+complexity, but this model's behavior meets at least four core criteria. One,
+as previously described, agent cognition and interactions cause behavior 
+to emerge in a bottom-up fashion. Two, because the model's macro-behavior derives from
+micro-interactions, one cannot causally separate the part from the whole. For
+example, suppose that State A increases its mechanization, and therefore neighboring
+State B also does so. One cannot say that State A \textit{caused} State B to
+raise its mechanization since State A's original increase itself derives from
+interaction with other neighbors, enemies, historical experiences, and the like. 
+Three, a positive feedback loop emerges. Four, system behavior is in some ways 
+resilient to variation in input parameter values. As shown in Table 6's parameter sweep, very
+different input values lead to about the same behavior: e.g., the same states
+are always relatively the most mechanized. That the model's behavior matches so
+many attributes of complex systems strengthens the belief that real-world
+mechanization is a complex system (put differently, that these findings are not
+just artifacts of a particular model).
 
 \subsection{Future Research}
+
+This paper's findings raise an obvious and important question: what is the
+missing ``brake'' that, if included in the model, 
+would hold mechanization to historical levels? Fully answering this question is
+beyond the scope of this paper.
 

### 2018-07-30 paper2/paper2.tex
foo

@@
 were not direct combatants to reform their own militaries
-\citep{scobell2011chinese}. There also is no guarantee that such cognition
+\citep{scobell2011chinese}. Similarly, Prussia sent military observers to the
+American Civil War and incorporated insights from there into its military
+doctrine \citep{scheibert2001prussian}. There also is no guarantee that such cognition
 follows typical assumptions of bounded rationality. 
@@
 would hold mechanization to historical levels? Fully answering this question is
-beyond the scope of this paper.
+beyond the scope of this paper. However, an extension to the base model
+provides insight that may guide future research.
+
+In the base model, the effects of an anti-mechanization lesson (defeat by an
+insurgency) only accrue to the state that suffered the defeat. In the extended
+model, the effect also can accrue to other states that were not direct participants 
+in the conflict. Put differently, agents can learn from the experiences of other
+states rather than solely from their own. This likely is a more accurate
+depiction of agent cognition because there is strong historical evidence of
+states revising their military force structure in response to to the perceived
+lessons of other states' wars (c.f. the discussion in section 2.2.2).
+
+This dynamic may be sufficient to curb model mechanization to historical levels
+by providing a strong countervailing positive feedback loop. In the base model,
+because only the direct losing side learns from an anti-mechanization lesson,
+the only way that
 
